### 2025-01-27 test.Rmd

---
title: "Geospatial"
author: 
  - Moritz Peist
  - Noemi Lucchi
  - Simon Vellin
date: "2025-01-27"
output: html_document
---

<hr>

```{css, echo=FALSE}
h1.title {
  text-align: center;
}

h4.author, h4.date {
  text-align: center;
}

.author {
  text-align: center;
}

.date {
  text-align: center;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<div style="text-align: center;">
  <img src="hw1/Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
</div>

<hr>

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.