### 2025-01-28 test.Rmd

---
title: "Geospatial"
author: 
  - Moritz Peist
  - Noemi Lucchi
  - Simon Vellin
date: "2025-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<table>
<tr>
<td><img src="hw1/data/paper3/c1960.jpg" width="400"/></td>
<td><img src="hw1/data/paper3/c2010.png" width="400"/></td>
</tr>
</table>