### 2025-01-29 hw2/hw2_MoritzPeistNoemiLucchiSimonVellin.Rmd
Initial submission for assignment 2 - TODO: Upload world data

---
title: "Geospatial - Assignment 2"
author: 
- Moritz Peist
- Noemi Lucchi
- Simon Vellin
output:
  html_document:
    toc: true
  pdf_document:
    toc: true
---

<hr>

```{css, echo=FALSE}
h1.title {
  text-align: center;
}

h4.author, h4.date {
  text-align: center;
}

.author {
  text-align: center;
}

.date {
  text-align: center;
}
```

```{r setup, include=FALSE}
# Disable warning messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div style="text-align: center;">
  <img src="../Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
</div>

<hr>

# Methodology & general code

In the following notebook we replicate the charts for all 5 papers. Contentwise, the code follows always the same structure. We first load the data, secondly manipulate it, then we create the chart and finally display it.

First we load the needed packages:

```{r}
# Main tidyverse (includes ggplot2)
library(tidyverse)

# Spatial packages
library(sf)
```

<hr>