### 2025-02-03 hw2/hw2_MoritzPeistNoemiLucchiSimonVellin.Rmd
Finished part 1

---
title: "Geospatial - Assignment 2"
author: 
- Moritz Peist
- Noemi Lucchi
- Simon Vellin
output:
  html_document:
    toc: true
  pdf_document:
    toc: true
---

<hr>

```{css, echo=FALSE}
h1.title {
  text-align: center;
}

h4.author, h4.date {
  text-align: center;
}

.author {
  text-align: center;
}

.date {
  text-align: center;
}
```

```{r setup, include=FALSE}
# Disable warning messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div style="text-align: center;">
  <img src="../Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
</div>

<hr>

# Methodology & general code

In this notebook, we combine various sources from [Natural Earth](https://www.naturalearthdata.com) to produce different types of plots. Throughout, we implement the methodologies recently discussed in class, such as distance computation, spatial merging, and filtering. Contentwise, we follow the assignment's instructions and provide the requested plots.

First we load the needed packages:

```{r}
# Main tidyverse (includes ggplot2)
library(tidyverse)

# Spatial packages
library(sf)
library(spData)
library(tidyverse)
library(units)
library(lwgeom)
```

<hr>

# Exercise 1: World and airports

We decided to use the [Natural Earth airports dataset](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/airports/) to plot the airports on a world map for this task. We downloaded all the data and load it from disk for speed and efficiency though a [package](https://cran.r-project.org/web/packages/rnaturalearth/vignettes/rnaturalearth.html) would exist to download the data directly.

## 1.1. Map of total population by country

### Data Loading

In order to plot the distribution of the total population by country, we import data from [Natural Earth Admin 0 - Countries](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/10m-admin-0-countries/) dataset which contains both the geometries of the countries and the corresponding population estimate.

```{r}
sf.world <- st_read("data/ne_10m_admin_0_countries", crs = 4326) %>%
  # Rename columns in the world dataset to make joining / operations easier
  rename(
    iso_a2 = ISO_A2,
    continent = CONTINENT
  )
```

We do not need any extensive data wrangling for this plot, as the data is already in a suitable format. Thus, we just go ahead and plot the total population by country.

### Plotting

```{r}
# Population in absolute terms
ggplot() +
  geom_sf(data = sf.world, aes(fill = POP_EST)) +
  scale_fill_gradient(name = "Population", low = "lightblue", high = "blue") +
  labs(title = "Total Population by Country", fill = "Population") +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

<hr>

## 1.2. Histogram of country population distribution by continent

### Data Loading

As for the previous plot, we only need the world data.

### Data Wrangling

First, we filter out countries without a continent and repair invalid polygons in case there are any. For a coarse-granular overview, we define population ranges by labeling countries based on their population size. After categorizing the countries into these ranges, we calculate the number of countries in each population range, grouped by continent. This gives us a summary of the distribution of countries by population size within each continent.
Finally, we ensure that the population ranges are ordered increasingly ("0-10M", "10M-50M", "50M-100M", "100M-500M", "500M+") by using the R factor datatype.

```{r}
# Remove Seven seas and Antarctica because they are NA
sf.world <- sf.world %>%
  filter(!is.na(continent)) %>%
  # Repair invalid polygons
  st_make_valid() %>%
  # Create the population ranges as factors
  mutate(pop_range = factor(
    case_when(
      POP_EST < 1e7 ~ "0-10M",
      POP_EST >= 1e7 & POP_EST < 5e7 ~ "10M-50M",
      POP_EST >= 5e7 & POP_EST < 1e8 ~ "50M-100M",
      POP_EST >= 1e8 & POP_EST < 5e8 ~ "100M-500M",
      TRUE ~ "500M+"
    ),
    levels = c("0-10M", "10M-50M", "50M-100M", "100M-500M", "500M+")
  ))

# Calculate the number of countries per range + country
population_distribution <- sf.world %>%
  group_by(continent, pop_range) %>%
  summarise(country_count = n(), .groups = "drop")
```

### Plotting

Now we start with a coarse granular plot to get an initial overview of the data.

```{r}
ggplot(population_distribution, aes(x = pop_range, y = country_count, fill = continent)) +
  geom_histogram(stat = "identity", position = "dodge") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Country Population Distribution by Continent", x = "Population Range",
    y = "Number of Countries", fill = "Continent"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

We plot the same data but with a more detailed view to better understand the distribution of countries by population size within each continent as asked by the task. However, we exclude Seven Seas and Antarctica as they contribute relatively little information. Moreover, we use a log-scale by population size, since country populations vary hugely, also by continent as this allows us to better visualize the data.

```{r}
sf.world_filt <- sf.world %>%
  mutate(
    continent = case_when(
      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
      TRUE ~ continent
    )
  ) %>%
  filter(!is.na(continent))

# Create a proper histogram using the continuous population data
ggplot(sf.world_filt, aes(x = POP_EST, fill = continent)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
  scale_x_log10(labels = scales::comma) + # Use log scale due to large population range
  facet_wrap(~continent, scales = "free_y") + # Separate panels for each continent
  labs(
    title = "Population Distribution by Continent",
    x = "Population (log scale)",
    y = "Count of Countries",
    fill = "Continent"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none", # Remove legend since we're using facets
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(
      angle = 45, # Rotate labels
      hjust = 1, # Align rotated labels
      size = 8 # Smaller font size
    )
  )
```

<hr>

## 1.3. Histogram of country-level average distances between locations and airports by continent.

### Data Loading

In addition to Natural Earth's population data, we incorporate also the datasets for [populated places (simple)](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/10m-populated-places/) (sf.pop) and [airports](https://www.naturalearthdata.com/downloads/10m-cultural-vectors/airports/) (sf.airports).

```{r}
sf.airports <- st_read("data/ne_10m_airports", crs = 4326)
sf.pop <- st_read("data/ne_10m_populated_places_simple", crs = 4326)
```

### Data Wrangling

This code computes and plots the average distances between populated places and airports across different continents, using two approaches: first, by computing the minimum distance for each city and averaging these values per country, and second, by directly averaging all city-airport distances within a country to allow for range-based visualization. Cities from the sf.pop dataset are joined with country and continent metadata from the sf.world dataset, while airports from sf.airports are joined with country metadata. Distances are calculated at two levels: in "distances" we compute the minimum distance from each city to the nearest airport and average these values per country, while in "distances_country" we directly average all city-airport distances within a country. Then we categorize "distances_country" into ranges (e.g., <50 km, <100 km) to ensure a better visualization. The first plot shows the distribution of city-level average distances by continent, while the second plot shows the number of countries in each distance range, grouped by continent.

**Important to note:** We only calculate the distance of the closest airport to each city per country. This means that we do not consider the distance of all populated places to all airports in a country, but only to the closest one. This is a simplification that allows us to reduce the number of distances to be computed and to avoid excessive granularity in the visualization, but also methodologically makes the most sense, otherwise distances in big countries will just be overboarding. Moreover, we add a buffer and use `st_nearest_feature` to avoid matching issues with coastal airports.

Next, we prepare the data for plotting the histograms of country-level average distances between locations and airports by continent.

```{r}
# 1. Prepare cities and airports with country metadata
cities_with_countries <- sf.world %>%
  st_drop_geometry() %>% # Drop geometry to avoid conflicts
  select(iso_a2, continent) %>% # Select relevant columns
  left_join( # Join with cities dataset
    sf.pop %>% select(name, iso_a2, ne_id, geometry),
    by = "iso_a2"
  ) %>%
  st_as_sf() %>% # Convert to sf object
  drop_na() # Remove rows with missing values

# 2. Create a small buffer around countries to catch coastal points
# Buffer size in degrees (since CRS is 4326)
buffer_size <- 0.01 # roughly 1km at the equator

airports_with_countries <- sf.airports %>%
  select(name, geometry) %>% # Select relevant columns
  # Join with buffered country polygons
  st_join(
    sf.world %>% # Join with country metadata
      select(iso_a2, continent) %>%
      st_buffer(dist = buffer_size),
    join = st_nearest_feature, # Use nearest feature instead of intersection
    left = TRUE # Keep all airports
  )

# 2. Calculate closest airport distance per city, then average per country
distances <- cities_with_countries %>%
  # For each country group:
  group_by(continent, iso_a2, ne_id) %>%
  summarise(
    avg_closest_distance_km = {
      current_iso <- cur_group()$iso_a2 # Capture current country code
      country_airports <- airports_with_countries %>%
        filter(iso_a2 == current_iso) # Filter airports for this country
      if (nrow(country_airports) == 0) {
        NA_real_ # Return NA if no airports exist
      } else {
        # Compute closest airport distance for each city in the country
        closest_distances <- st_distance(
          geometry, # City geometries
          country_airports # Airports in the current country
        ) %>%
          apply(1, min) %>% # Find minimum distance per city (row)
          as.numeric() / 1000 # Convert meters to kilometers
        mean(closest_distances, na.rm = TRUE)
      }
    }
  ) %>%
  ungroup() %>% # Remove grouping
  drop_na() # Remove countries with no airports
```

### Plotting

Now, we plot the histogram of country level average distances between locations and airports by continent - as asked for by the exercise. Once again we filter out the Seven Seas and Antarctica as they do not contribute much information and since only 16 airports are impacted by this anyways.

```{r}
distances_filt <- distances %>%
  mutate(
    continent = case_when(
      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
      TRUE ~ continent
    )
  ) %>%
  filter(!is.na(continent))

# Plot 1
ggplot(distances_filt, aes(x = avg_closest_distance_km, fill = continent)) +
  geom_histogram(bins = 30, position = "dodge") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Average Distance Between Locations and Airports by Continent",
    x = "Average Distance to nearest Airport (km)", y = "Number of Populated Places", fill = "Continent"
  ) +
  facet_wrap(~continent, scales = "free") +
  theme_minimal() +
  theme(
    legend.position = "none", # Remove legend since we're using facets
    plot.title = element_text(hjust = 0.5)
  )
```

What becomes immediately apparent is that most of the continents have some outliers with huge distances to the nearest airport. This is especially true for Africa and Asia, where some cities are very remote and far away from airports, but also for Europe, where some cities are located on islands or countries like Russia may have very large distances to the next airport. The most extreme example is Uelene. A russian village just south of the arctic cycle. In this sense, the populated places dataset might be misleading as it also includes also very remote places which are sparse in population and far away from airports. Nevertheless, also our join by `st_nearest_feature()` might mismatch some airports and cities, especially in coastal areas.

Since we have these issues, let's look at the airports also in a more coarse-granular way to get a better understanding of the data (once again we exclude Seven Seas and Antarctica).

```{r}
# 3. Calculate average distances at country level. With this approach, instead of
# computing the closest airport distance per city and then averaging per country,
# we compute all the distances within a country and then we average them.
distances_country <- cities_with_countries %>%
  group_by(continent, iso_a2) %>% # Removed ne_id to group at country level
  summarise(
    avg_closest_distance_km = {
      current_iso <- cur_group()$iso_a2
      country_airports <- airports_with_countries %>%
        filter(iso_a2 == current_iso)
      if (nrow(country_airports) == 0) {
        NA_real_
      } else {
        closest_distances <- st_distance(
          geometry,
          country_airports
        ) %>%
          apply(1, min) %>%
          as.numeric() / 1000
        mean(closest_distances, na.rm = TRUE)
      }
    },
    country_name = first(name)
  ) %>%
  ungroup() %>%
  drop_na()

# 4. Create distance ranges – The previous step ensures a reasonable number of ranges.
# If we had used the distances computed earlier, the higher granularity would have resulted in an excessive number of ranges, leading to a poor visualization quality.
distances_ranges <- distances_country %>%
  mutate(
    distance_range = cut(avg_closest_distance_km,
      breaks = seq(0, max(avg_closest_distance_km, na.rm = TRUE) + 50, by = 50),
      labels = paste0("<", seq(50, max(avg_closest_distance_km, na.rm = TRUE) + 50, by = 50), "km"),
      include.lowest = TRUE
    ),
    continent = case_when(
      continent %in% c("Seven seas (open ocean)", "Antarctica") ~ NA_character_,
      TRUE ~ continent
    )
  ) %>%
  filter(!is.na(continent))
```

Let's also plot the coarse-granular view of the data to get a better understanding of the distribution of average distances between locations and airports by continent.

```{r}
ggplot(distances_ranges, aes(x = distance_range, fill = continent)) +
  geom_bar(position = "stack") +
  scale_fill_brewer(palette = "Set2") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Average Distance Between Locations and Airport by Ranges",
    x = "Average Distance in Ranges", y = "Number of Countries", fill = "continent"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

<hr>

# Exercise 2: Porteuos data