### 2025-03-15 final/final_MoritzPeistNoemiLucchiSimonVellin.Rmd
Initial commit final assignment

---
title: "R Notebook"
output: html_notebook
---

Load packages

```{r Packages}
library(data.table)   # For faster data manipulation
library(tidyverse)    # For data manipulation and visualization
library(sf)           # For spatial data handling
library(leaflet)      # For interactive maps
library(leaflet.extras) # For additional leaflet features
library(mapview)      # For easier map visualization
library(tmap)         # For thematic maps
library(tigris)       # For US road networks
library(future)       # For parallel processing
library(future.apply) # For parallel processing with apply functions

# Create directories if they don't exist
if (!dir.exists("./data/tigris")) {
  dir.create("./data/tigris", recursive = TRUE)
}

# Set custom cache directory (optional)
options(tigris_cache_dir = "./data/tigris")
# Configure tigris to use caching
options(tigris_use_cache = TRUE)
```

# Set-up convenience functions for parallel processing (as data is quite large)
```{r Parallel Buffering}
# Function to create buffers in batches with proper projection
create_buffers_in_batches <- function(sf_object, buffer_dist, batch_size = 500) {
  # First, reproject to an appropriate projected CRS for the region
  # For California, UTM Zone 10N (EPSG:26910) or 11N (EPSG:26911) works well
  # If we're unsure about your region, a web mercator projection (3857) is a reasonable default
  message("Reprojecting data to a meter-based CRS...")
  sf_object_projected <- st_transform(sf_object, 3310)  # Web Mercator

  n_features <- nrow(sf_object_projected)
  n_batches <- ceiling(n_features / batch_size)

  # Create empty list to store results
  buffer_list <- vector("list", n_batches)

  # Process in batches
  for (i in 1:n_batches) {
    start_idx <- (i-1) * batch_size + 1
    end_idx <- min(i * batch_size, n_features)

    cat(sprintf("Processing batch %d of %d (features %d to %d)\n", 
                i, n_batches, start_idx, end_idx))

    # Extract batch
    batch <- sf_object_projected[start_idx:end_idx, ]

    # Create buffer (with parallel processing within sf)
    buffer_list[[i]] <- st_buffer(batch, dist = buffer_dist)
  }

  # Combine results
  result <- do.call(rbind, buffer_list)

  # Reproject back to original CRS if needed
  message("Reprojecting results back to original CRS...")
  st_transform(result, st_crs(sf_object))
}
```

Load data (filter for California for now to limit data set size)

```{r Data Load Accidents}
# Efficient approach
df.acc <- fread("data/us_accidents/US_accidents_March23.csv")[
  # Filter date range to 2016-2021
  as.Date(Start_Time) >= as.Date("2016-01-01") & 
  as.Date(Start_Time) <= as.Date("2021-12-31") & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = data.table::year(Start_Time),
  quarter = data.table::quarter(Start_Time),
  month = data.table::month(Start_Time),
  # Calculate duration (assuming End_Time exists in the dataset)
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()  # Convert to tibble only at the end for performance
```

```{r Data Load Constructions}
df.const <- fread("data/us_constructions/US_constructions_Dec21.csv")[
  # Filter date range to 2016-2021
  as.Date(Start_Time) >= as.Date("2016-01-01") & 
  as.Date(Start_Time) <= as.Date("2021-12-31") & 
    # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = year(Start_Time),
  quarter = quarter(Start_Time),
  month = month(Start_Time),
  # Calculate duration (assuming End_Time exists in the dataset)
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()  # Convert to tibble only at the end for performance
```

Convert dataframes to sf objects

```{r Geospatial Data}
# Convert accident data to sf object
df.acc.sf <- df.acc %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)

# Convert construction data to sf object
df.const.sf <- df.const %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)
```

```{r Data Load Roads}
# Get US roads (this can be slow for the entire US, so we might want to limit by state)

# First get all counties in California
ca_counties <- counties("CA", year = 2021)

# Then get roads for each county
ca_roads_list <- lapply(ca_counties$COUNTYFP, function(county_code) {
  roads("CA", county = county_code, year = 2021)
})

# Optionally combine all county road data into one object
ca_roads <- do.call(rbind, ca_roads_list)
```

```{r Static Maps}
# Static map for Accidents with year coloring
# First, store your plot in a variable
ca_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.acc.sf %>% filter(State == "CA"), 
          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
  theme_minimal() +
  labs(title = "US Road Accidents by Year (2016-2021)") +
  theme(legend.position = "bottom")
print(ca_plot)
# Then save it using ggsave
#ggsave(filename = "imgs/california_accidents.png", 
#       plot = ca_plot,
#       width = 10, # width in inches
#       height = 8, # height in inches
#       dpi = 300)  # resolution

# Construction sites map with year coloring
# First, store our plot in a variable
ca_const_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.const.sf %>% filter(State == "CA"), 
          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
  theme_minimal() +
  labs(title = "US Construction Sites by Year (2016-2021)") +
  theme(legend.position = "bottom")
print(ca_const_plot)
# Then save it using ggsave
#ggsave(filename = "imgs/california_construction.png", 
#       plot = ca_const_plot,
#       width = 10, # width in inches
#       height = 8, # height in inches
#       dpi = 300)  # resolution
```

## Heatmaps - interactive

```{r Interactive Maps I}
# For interactive maps (often better for large datasets)
# Accidents map
accident_map <- leaflet() %>%
  addTiles() %>%
  addHeatmap(data = df.acc.sf, 
             intensity = ~1,
             radius = 8, 
             blur = 10) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>% # Center to CA
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display maps
accident_map
```

```{r Interactive Maps II}
# Construction sites map with California focus and bounds constraints
construction_map <- leaflet() %>%
  addTiles() %>%
  addHeatmap(data = df.const.sf, 
             intensity = ~1,
             radius = 8, 
             blur = 10,
             gradient = c("yellow", "orange", "red")) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display the map
construction_map
```

## Buffer zone analysis

```{r Buffer Zone Analysis}
# Set up parallel processing
future::plan(future::multisession, workers = parallel::detectCores() - 1)

# Set sf to use parallel processing
sf_use_s2(FALSE)  # Disable S2 spherical geometry
cores <- parallel::detectCores() - 1

# Process each dataset with batching
cat("Creating point buffers...\n")
const_buffers <- create_buffers_in_batches(df.const.sf, buffer_dist = 1000, batch_size = 500)

# Return to sequential processing
future::plan(future::sequential)
sf_use_s2(TRUE)  # Re-enable S2 spherical geometry

# Find accidents within construction zones
# Use a spatial index and parallel processing
plan(multisession, workers = parallel::detectCores() - 1)  # Use parallel processing

accidents_in_construction <- df.acc.sf %>%
  st_join(ca_const_buffers, join = st_within, prepared = TRUE)

# Reset to sequential processing when done
plan(sequential)

# Count accidents per construction zone
accidents_per_zone <- accidents_in_construction %>%
  group_by(ID) %>%  # Grouping by the ID column
  summarize(accident_count = n())

# Visualize construction zones with accident counts
# First, store our plot in a variable
accident_zones_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = accidents_per_zone, aes(fill = accident_count), alpha = 0.7) +
  scale_fill_viridis_c() +
  theme_minimal() +
  labs(title = "Accident Counts within Construction Zones (2016-2021)")

print(accident_zones_plot)

# Then save it using ggsave
#ggsave(filename = "imgs/accident_counts_construction_zones.png", 
#       plot = accident_zones_plot,
#       width = 10, # width in inches
#       height = 8, # height in inches
#       dpi = 300)  # resolution
```

Advanced construction sites:

```{r SF linestring}
# Convert construction data to linestring sf object
df.const.lines <- df.const %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng) & !is.na(End_Lat) & !is.na(End_Lng)) %>%
  mutate(geometry = pmap(list(Start_Lng, Start_Lat, End_Lng, End_Lat), 
                         function(start_lng, start_lat, end_lng, end_lat) {
                           coords <- matrix(c(start_lng, start_lat, 
                                             end_lng, end_lat), 
                                           ncol = 2, byrow = TRUE)
                           st_linestring(coords)
                         })) %>%
  st_sf(crs = 4326)
```

```{r Maps with Construction Lines}
# Static map with construction line segments
ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.const.lines %>% filter(State == "CA"), 
          color = "orange", size = 1) +
  theme_minimal() +
  labs(title = "US Construction Segments (2016-2021)")

# Interactive map with construction segments focused on California
construction_segment_map <- leaflet() %>%
  addTiles() %>%
  addPolylines(data = df.const.lines, 
               color = "orange",
               weight = 3,
               opacity = 0.7,
               popup = ~paste("ID:", ID, "<br>Duration:", Duration)) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display the map
construction_segment_map
```

```{r Construction Impact Analysis with Buffers}
# Set up parallel processing
future::plan(future::multisession, workers = parallel::detectCores() - 1)

# Set sf to use parallel processing
sf_use_s2(FALSE)  # Disable S2 spherical geometry
cores <- parallel::detectCores() - 1

cat("Creating line segment buffers...\n")
const_segment_buffers <- create_buffers_in_batches(df.const.lines, buffer_dist = 1000, batch_size = 250)

# Return to sequential processing
future::plan(future::sequential)
sf_use_s2(TRUE)  # Re-enable S2 spherical geometry

# Find accidents near construction segments
accidents_near_construction <- df.acc.sf %>%
  st_join(const_segment_buffers, join = st_intersects)

# Analyze accident rate per construction segment length
accidents_per_segment <- accidents_near_construction %>%
  group_by(ID) %>%  # Assuming there's an ID column for construction segments
  summarize(
    accident_count = n(),
    segment_length = as.numeric(st_length(geometry[1])),  # Length in meters
    accidents_per_km = accident_count / (segment_length/1000)
  )

# Visualize segments by accident rate
ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = accidents_per_segment, aes(color = accidents_per_km), size = 1.5) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Accident Rate per Construction Segment (2016-2021)",
       color = "Accidents/km")
```

```{r Temporal Analysis and Severity Comparison with Buffers}
# Temporal analysis - construction impact over time
accidents_by_time <- accidents_near_construction %>%
  mutate(month = floor_date(as.Date(Start_Time), "month")) %>%
  group_by(month) %>%
  summarize(accident_count = n())

# Compare accident severity near vs. away from construction
severity_comparison <- bind_rows(
  accidents_near_construction %>% 
    mutate(location = "Near construction") %>% 
    select(Severity, location),

  anti_join(df.acc.sf, accidents_near_construction, by = "ID") %>% 
    mutate(location = "Away from construction") %>% 
    select(Severity, location)
)

# Visualize severity comparison
ggplot(severity_comparison, aes(x = Severity, fill = location)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(title = "Accident Severity: Near vs. Away from Construction",
       x = "Severity Level",
       y = "Count",
       fill = "Location")
```