### 2025-03-18 final/streamlined_nb.Rmd

---
title: "Geospatial Data Science - Final project"
author:
- Moritz Peist
- Noemi Lucchi
- Simon Vellin
subtitle: Analysis of US road accidents and construction sites
output:
  html_notebook:
    df_print: paged
    toc: true
  html_document:
    toc: true
    df_print: paged
---

<hr>

```{css, echo=FALSE}
h1.title {
  text-align: center;
}

h4.author, h4.date {
  text-align: center;
}

.author {
  text-align: center;
}

.date {
  text-align: center;
}

/* Add styling for subtitle */
h3.subtitle {
  text-align: center;
  font-style: italic;
  margin-top: -10px;
  color: #606060;
}
```

<div style="text-align: center;">
  <img src="../Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
</div>

<hr>

## 1. Setup and Packages

```{r setup, message=FALSE, warning=FALSE}
# Load all necessary packages
library(data.table)   # For faster data manipulation
library(tidyverse)    # For data manipulation and visualization
library(sf)           # For spatial data handling
library(leaflet)      # For interactive maps
library(leaflet.extras) # For additional leaflet features
library(mapview)      # For easier map visualization
library(tmap)         # For thematic maps
library(tigris)       # For US road networks
library(future)       # For parallel processing
library(future.apply) # For parallel processing with apply functions
library(fixest)       # For efficient fixed effects estimation
library(modelsummary) # For regression tables
library(patchwork)    # For combining plots
library(MatchIt)      # For matching
library(lubridate)    # For date handling

# Create directories if they don't exist
if (!dir.exists("./data/tigris")) {
  dir.create("./data/tigris", recursive = TRUE)
}
if (!dir.exists("./data/processed")) {
  dir.create("./data/processed", recursive = TRUE)
}
if (!dir.exists("./imgs")) {
  dir.create("./imgs", recursive = TRUE)
}

# Set custom cache directory
options(tigris_cache_dir = "./data/tigris")
# Configure tigris to use caching
options(tigris_use_cache = TRUE)
```

## 2. Helper Functions for Spatial Operations

```{r helper_functions}
# Function to create buffers in batches with proper projection
create_buffers_in_batches <- function(sf_object, buffer_dist, batch_size = 500) {
  message("Reprojecting data to a meter-based CRS...")
  sf_object_projected <- st_transform(sf_object, 3310)  # California Albers

  n_features <- nrow(sf_object_projected)
  n_batches <- ceiling(n_features / batch_size)

  # Create empty list to store results
  buffer_list <- vector("list", n_batches)

  # Process in batches
  for (i in 1:n_batches) {
    start_idx <- (i-1) * batch_size + 1
    end_idx <- min(i * batch_size, n_features)

    cat(sprintf("Processing batch %d of %d (features %d to %d)\n", 
                i, n_batches, start_idx, end_idx))

    # Extract batch
    batch <- sf_object_projected[start_idx:end_idx, ]

    # Create buffer
    buffer_list[[i]] <- st_buffer(batch, dist = buffer_dist)
  }

  # Combine results
  result <- do.call(rbind, buffer_list)

  # Reproject back to original CRS
  message("Reprojecting results back to original CRS...")
  st_transform(result, st_crs(sf_object))
}

# Unified function to create both point and line buffers
create_all_buffers <- function(point_sf, line_sf, buffer_dist = 300) {
  # Process point buffers
  point_buffers <- point_sf %>%
    mutate(construction_id = if("ID" %in% names(.)) ID else row_number(),
           construction_type = "point") %>%
    select(construction_id, construction_type, Start_Time, End_Time, geometry) %>%
    rename(start_time = Start_Time, end_time = End_Time) %>%
    create_buffers_in_batches(buffer_dist = buffer_dist, batch_size = 500)

  # Process line buffers if provided
  if(!missing(line_sf) && nrow(line_sf) > 0) {
    # Calculate segment length once
    line_sf_with_length <- line_sf %>%
      st_transform(3310) %>%
      mutate(segment_length_m = as.numeric(st_length(geometry))) %>%
      st_transform(4326)

    line_buffers <- line_sf_with_length %>%
      mutate(construction_id = if("ID" %in% names(.)) ID else row_number() + nrow(point_sf),
             construction_type = "line") %>%
      select(construction_id, construction_type, Start_Time, End_Time, 
             segment_length_m, geometry) %>%
      rename(start_time = Start_Time, end_time = End_Time) %>%
      create_buffers_in_batches(buffer_dist = buffer_dist, batch_size = 250)

    # Combine buffers
    return(list(
      combined_buffers = bind_rows(
        select(point_buffers, -construction_type),
        select(line_buffers, -construction_type, -segment_length_m)
      ),
      point_buffers = point_buffers,
      line_buffers = line_buffers,
      line_data = line_sf_with_length
    ))
  }

  # Return just point buffers if no lines
  return(list(
    combined_buffers = point_buffers,
    point_buffers = point_buffers
  ))
}

# Helper function to find accidents near constructions
find_accidents_near_construction <- function(accident_sf, buffer_results) {
  # For point buffers
  point_spatial_matches <- accident_sf %>%
    mutate(accident_id = if("ID" %in% names(.)) ID else row_number(),
           accident_severity = if("Severity" %in% names(.)) Severity else NA_integer_,
           accident_time = Start_Time) %>%
    select(accident_id, accident_severity, accident_time, geometry) %>%
    st_join(buffer_results$point_buffers, join = st_within) %>%
    filter(!is.na(construction_id) & 
           accident_time >= start_time & 
           accident_time <= end_time) %>%
    select(accident_id, accident_severity, accident_time, construction_id, construction_type)

  # For line buffers if they exist
  if ("line_buffers" %in% names(buffer_results)) {
    line_spatial_matches <- accident_sf %>%
      mutate(accident_id = if("ID" %in% names(.)) ID else row_number(),
             accident_severity = if("Severity" %in% names(.)) Severity else NA_integer_,
             accident_time = Start_Time) %>%
      select(accident_id, accident_severity, accident_time, geometry) %>%
      st_join(buffer_results$line_buffers, join = st_within) %>%
      filter(!is.na(construction_id) & 
             accident_time >= start_time & 
             accident_time <= end_time) %>%
      select(accident_id, accident_severity, accident_time, construction_id, 
             construction_type, segment_length_m)

    # Combine both types of matches
    return(list(
      point_matches = point_spatial_matches,
      line_matches = line_spatial_matches,
      all_matches = bind_rows(
        point_spatial_matches,
        select(line_spatial_matches, -segment_length_m)
      )
    ))
  }

  # Return just point matches if no line buffers
  return(list(
    point_matches = point_spatial_matches,
    all_matches = point_spatial_matches
  ))
}

# Helper function to analyze accidents near construction
analyze_accidents_near_construction <- function(spatial_matches, accident_sf, line_data) {
  # Count construction sites per accident
  construction_counts <- spatial_matches$all_matches %>%
    st_set_geometry(NULL) %>%
    group_by(accident_id) %>%
    summarize(nearby_construction_count = n())

  # Randomly select one construction site per accident
  accident_construction_analysis <- spatial_matches$all_matches %>%
    st_set_geometry(NULL) %>%
    left_join(construction_counts, by = "accident_id") %>%
    mutate(
      temporal_overlap = 1,
      is_near_active_construction = 1
    ) %>%
    group_by(accident_id) %>%
    slice_sample(n = 1) %>%
    ungroup()

  # Calculate accident rates for line segments
  if ("line_matches" %in% names(spatial_matches) && !is.null(line_data)) {
    construction_segment_accident_rates <- line_data %>%
      mutate(construction_id = if("ID" %in% names(.)) ID else row_number()) %>%
      left_join(
        spatial_matches$line_matches %>%
          st_set_geometry(NULL) %>%
          group_by(construction_id) %>%
          summarize(accident_count = n()),
        by = "construction_id"
      ) %>%
      mutate(
        accident_count = replace_na(accident_count, 0),
        accidents_per_km = accident_count / (segment_length_m / 1000)
      )
  } else {
    construction_segment_accident_rates <- NULL
  }

  # Add construction info to all accidents
  accidents_with_construction <- accident_sf %>%
    mutate(accident_id = if("ID" %in% names(.)) ID else row_number()) %>%
    left_join(
      accident_construction_analysis %>%
        select(accident_id, construction_id, nearby_construction_count, construction_type),
      by = "accident_id"
    ) %>%
    mutate(near_construction = ifelse(is.na(construction_id), 0, 1))

  # Return results
  return(list(
    accident_construction_analysis = accident_construction_analysis,
    construction_segment_accident_rates = construction_segment_accident_rates,
    accidents_with_construction = accidents_with_construction,
    total_accidents = nrow(accident_sf),
    accidents_near_construction = nrow(accident_construction_analysis),
    pct_near_construction = round(nrow(accident_construction_analysis)/nrow(accident_sf)*100, 2)
  ))
}

# NEW FUNCTION: Create analysis panel with improved spatial matching
create_analysis_panel <- function(df.const.sf, df.acc.sf, ca_roads, buffer_dist = 300) {
  # Step 1: Create a more appropriate spatial unit - larger road chunks
  # Dissolve/merge road segments by road name and type to create longer segments
  cat("Simplifying road network...\n")
  simplified_roads <- ca_roads %>%
    group_by(FULLNAME, RTTYP) %>%
    summarize(geometry = st_union(geometry), .groups = "drop") %>%
    mutate(road_id = row_number(),
           road_length_km = as.numeric(st_length(st_transform(geometry, 3310)))/1000)

  # Step 2: Create construction buffers and find which roads they intersect
  cat("Creating construction buffers and finding affected roads...\n")
  const_buffers <- st_buffer(st_transform(df.const.sf, 3310), buffer_dist) %>%
    st_transform(st_crs(ca_roads))

  road_construction <- simplified_roads %>%
    st_join(const_buffers %>% 
              select(ID, Start_Time, End_Time), join = st_intersects) %>%
    filter(!is.na(ID)) %>%
    mutate(
      const_start = floor_date(Start_Time, "month"),
      const_end = floor_date(End_Time, "month")
    ) %>%
    st_set_geometry(NULL) %>%
    select(road_id, FULLNAME, RTTYP, const_start, const_end, ID)

  # Step 3: Create a panel with monthly observations for all roads
  cat("Creating panel dataset...\n")
  months <- seq(as.Date("2021-01-01"), as.Date("2021-12-01"), by = "month")

  # Create base panel with all road-month combinations
  panel_base <- expand.grid(
    road_id = simplified_roads$road_id,
    month = months
  ) %>% as_tibble()

  # Add construction activity flags
  cat("Adding construction treatment indicators...\n")
  panel_with_construction <- panel_base %>%
    left_join(
      road_construction %>%
        # Expand each construction to all months it spans
        group_by(road_id, ID) %>%
        do({
          months_seq <- seq(from = .$const_start[1], to = .$const_end[1], by = "month")
          data.frame(
            road_id = .$road_id[1],
            month = months_seq,
            has_construction = 1,
            const_id = .$ID[1]
          )
        }) %>%
        ungroup() %>%
        # Consolidate multiple constructions on same road-month
        group_by(road_id, month) %>%
        summarize(
          has_construction = 1,
          construction_count = n(),
          .groups = "drop"
        ),
      by = c("road_id", "month")
    ) %>%
    mutate(has_construction = ifelse(is.na(has_construction), 0, has_construction))

  # Step 4: Count accidents for each road segment by month
  cat("Matching accidents to roads...\n")
  # Buffer roads slightly to catch nearby accidents
  road_buffers <- st_buffer(st_transform(simplified_roads, 3310), 50) %>%
    st_transform(st_crs(df.acc.sf))

  accident_counts <- df.acc.sf %>%
    mutate(month = floor_date(Start_Time, "month")) %>%
    st_join(road_buffers %>% select(road_id), join = st_intersects) %>%
    filter(!is.na(road_id)) %>%
    st_set_geometry(NULL) %>%
    group_by(road_id, month) %>%
    summarize(
      accident_count = n(),
      severe_accident_count = sum(Severity >= 3, na.rm = TRUE),
      avg_severity = mean(Severity, na.rm = TRUE),
      .groups = "drop"
    )

  # Step 5: Create final panel with all variables
  cat("Finalizing panel dataset...\n")
  panel_final <- panel_with_construction %>%
    left_join(accident_counts, by = c("road_id", "month")) %>%
    left_join(
      simplified_roads %>%
        st_set_geometry(NULL) %>%
        select(road_id, road_length_km, FULLNAME, RTTYP),
      by = "road_id"
    ) %>%
    # Add road type classification
    mutate(
      road_type = case_when(
        grepl("I", RTTYP) ~ "Interstate",
        grepl("U", RTTYP) ~ "US Highway",
        grepl("S", RTTYP) ~ "State Highway",
        TRUE ~ "Other"
      ),
      # Fill in missing accident counts with 0
      accident_count = replace_na(accident_count, 0),
      severe_accident_count = replace_na(severe_accident_count, 0),
      avg_severity = replace_na(avg_severity, 0),
      # Calculate per-km rates
      accidents_per_km = accident_count / road_length_km,
      severe_accidents_per_km = severe_accident_count / road_length_km,
      # Extract time variables
      year = year(month),
      month_num = month(month),
      is_winter = month_num %in% c(12, 1, 2),
      is_summer = month_num %in% c(6, 7, 8)
    )

  # Add urban classification
  cat("Adding urban/rural classification...\n")
  urban_areas <- urban_areas(year = 2021)

  # Check which roads intersect urban areas
  urban_roads <- simplified_roads %>%
    st_transform(st_crs(urban_areas)) %>%
    st_join(urban_areas %>% select(UACE10, NAME10, UATYP10), join = st_intersects) %>%
    mutate(
      is_urban = !is.na(UACE10),
      urban_type = ifelse(is.na(UATYP10), "Rural", UATYP10)
    ) %>%
    st_set_geometry(NULL) %>%
    select(road_id, is_urban, urban_type)

  # Join urban classification
  panel_final <- panel_final %>%
    left_join(urban_roads, by = "road_id")

  # Add weather data if available (average by month for simplicity)
  if ("Temperature(F)" %in% names(df.acc.sf) && "Precipitation(in)" %in% names(df.acc.sf)) {
    cat("Adding weather data...\n")
    weather_data <- df.acc.sf %>%
      mutate(month = floor_date(Start_Time, "month")) %>%
      group_by(month) %>%
      summarize(
        avg_temperature = mean(`Temperature(F)`, na.rm = TRUE),
        avg_precipitation = mean(`Precipitation(in)`, na.rm = TRUE),
        pct_adverse_weather = mean(Weather_Condition %in% 
                                  c("Rain", "Snow", "Fog", "Thunderstorm"), na.rm = TRUE),
        .groups = "drop"
      ) %>%
      st_set_geometry(NULL)

    panel_final <- panel_final %>%
      left_join(weather_data, by = "month")
  }

  # Print diagnostic information
  cat("\nSummary Statistics:\n")
  cat("Total unique roads:", nrow(simplified_roads), "\n")
  cat("Roads with construction:", length(unique(road_construction$road_id)), "\n")
  cat("Road-months with construction:", sum(panel_final$has_construction), "\n")
  cat("Roads with accidents:", length(unique(accident_counts$road_id)), "\n")
  cat("Total accidents matched to roads:", sum(accident_counts$accident_count), "\n")
  cat("Final panel dimensions:", nrow(panel_final), "x", ncol(panel_final), "\n")

  return(list(
    panel_data = panel_final,
    simplified_roads = simplified_roads,
    road_construction = road_construction,
    accident_counts = accident_counts
  ))
}
```

## 3. Regression and Robustness Functions

```{r regression_functions}
# Run all main regression models
run_regression_models <- function(panel_data) {
  # 1. Base DiD Model
  model1 <- feols(
    accident_count ~ has_construction | road_id + month, 
    data = panel_data
  )

  # 2. Extended Model with Urban/Rural Heterogeneity
  model2 <- feols(
    accident_count ~ has_construction + has_construction:is_urban | road_id + month, 
    data = panel_data
  )

  # 3. Full Model with additional controls
  # Check if weather data is available
  if (all(c("avg_temperature", "avg_precipitation", "pct_adverse_weather") %in% names(panel_data))) {
    model3 <- feols(
      accident_count ~ has_construction + has_construction:is_urban + 
                      has_construction:is_winter + has_construction:is_summer +
                      avg_temperature + avg_precipitation + pct_adverse_weather | 
                      road_id + month,
      data = panel_data
    )
  } else {
    model3 <- feols(
      accident_count ~ has_construction + has_construction:is_urban + 
                      has_construction:is_winter + has_construction:is_summer | 
                      road_id + month,
      data = panel_data
    )
  }

  # 4. Accident rate model (normalized by segment length)
  model4 <- feols(
    accidents_per_km ~ has_construction + has_construction:is_urban | road_id + month, 
    data = panel_data
  )

  # 5. Severity model
  model5 <- feols(
    avg_severity ~ has_construction + has_construction:is_urban | road_id + month, 
    data = panel_data
  )

  return(list(
    base_model = model1,
    urban_het_model = model2,
    full_model = model3,
    rate_model = model4,
    severity_model = model5,
    all_models = list(model1, model2, model3, model4, model5)
  ))
}

# Helper functions for robustness checks
create_event_study_data <- function(panel_data) {
  panel_data %>%
    group_by(road_id) %>%
    mutate(
      first_construction = if(any(has_construction == 1)) min(month[has_construction == 1]) else NA,
      rel_month = as.numeric(difftime(month, first_construction, units = "days")) / 30,
      rel_period = case_when(
        is.na(rel_month) ~ NA_character_,
        rel_month <= -3 ~ "t-3+",
        rel_month <= -2 ~ "t-2",
        rel_month <= -1 ~ "t-1",
        rel_month < 0 ~ "t-0",
        rel_month < 1 ~ "t+0",
        rel_month < 2 ~ "t+1",
        rel_month < 3 ~ "t+2",
        TRUE ~ "t+3+"
      )
    ) %>%
    ungroup() %>%
    mutate(
      rel_period = factor(rel_period, 
                        levels = c("t-3+", "t-2", "t-1", "t-0", "t+0", "t+1", "t+2", "t+3+"))
    )
}

run_placebo_tests <- function(panel_data, iterations = 100) {
  placebo_results <- data.frame(iteration = 1:iterations, estimate = NA, std.error = NA)
  for(i in 1:iterations) {
    set.seed(i)
    temp_data <- panel_data %>%
      mutate(placebo_construction = sample(has_construction))

    temp_model <- feols(
      accident_count ~ placebo_construction | road_id + month, 
      data = temp_data
    )

    placebo_results$estimate[i] <- coef(temp_model)[1]
    placebo_results$std.error[i] <- se(temp_model)[1]
  }
  return(placebo_results)
}

run_fe_variations <- function(panel_data) {
  list(
    "No FE" = feols(accident_count ~ has_construction, data = panel_data),
    "Road FE" = feols(accident_count ~ has_construction | road_id, data = panel_data),
    "Time FE" = feols(accident_count ~ has_construction | month, data = panel_data),
    "Both FE" = feols(accident_count ~ has_construction | road_id + month, data = panel_data)
  )
}

run_transform_variations <- function(panel_data) {
  list(
    "Linear" = feols(accident_count ~ has_construction | road_id + month, data = panel_data),
    "Log" = feols(log(accident_count + 1) ~ has_construction | road_id + month, data = panel_data),
    "Poisson" = fepois(accident_count ~ has_construction | road_id + month, data = panel_data)
  )
}

# Run all robustness checks
run_robustness_checks <- function(panel_data, base_model) {
  # Results list
  robustness_results <- list()

  # 1. Event study approach
  panel_data_event <- create_event_study_data(panel_data)
  event_model <- feols(
    accident_count ~ i(rel_period, ref = "t-1") | road_id + month,
    data = panel_data_event %>% filter(!is.na(rel_period))
  )
  robustness_results$event_model <- event_model
  robustness_results$panel_data_event <- panel_data_event

  # 2. Placebo tests
  placebo_results <- run_placebo_tests(panel_data, 100)
  robustness_results$placebo_results <- placebo_results

  # 3. Fixed effects variations
  fe_models <- run_fe_variations(panel_data)
  robustness_results$fe_models <- fe_models

  # 4. Transform outcome variables
  transform_models <- run_transform_variations(panel_data)
  robustness_results$transform_models <- transform_models

  # 5. Add matching (simplified, assumes is_urban and road_type are factors)
  matched_data <- matchit(
    has_construction ~ road_type + road_length_km + is_urban,
    data = panel_data %>% 
      filter(month == min(month)) %>%
      mutate(
        road_type = factor(road_type),
        is_urban = factor(is_urban)
      ),
    method = "nearest"
  )
  robustness_results$matched_data <- matched_data

  # Return all results
  return(robustness_results)
}
```

## 4. Visualization Functions

```{r vis_functions}
# Create static maps
create_static_maps <- function(df.acc.sf, df.const.sf, df.const.lines, ca_roads) {
  # Accident map
  ca_plot <- ggplot() +
    geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
    geom_sf(data = df.acc.sf, 
            aes(color = factor(year)), alpha = 0.7, size = 0.5) +
    scale_color_viridis_d(name = "Year") +
    theme_minimal() +
    labs(title = "California Road Accidents (2021)") +
    theme(legend.position = "bottom")

  # Construction map
  ca_const_plot <- ggplot() +
    geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
    geom_sf(data = df.const.sf, 
            aes(color = factor(year)), alpha = 0.7, size = 0.5) +
    scale_color_viridis_d(name = "Year") +
    theme_minimal() +
    labs(title = "California Construction Sites (2021)") +
    theme(legend.position = "bottom")

  # Construction line segments map
  lines_plot <- ggplot() +
    geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
    geom_sf(data = df.const.lines, 
            color = "orange", size = 1) +
    theme_minimal() +
    labs(title = "California Construction Segments (2021)")

  return(list(
    accidents_map = ca_plot,
    construction_map = ca_const_plot,
    construction_lines_map = lines_plot
  ))
}

# Create heatmaps
create_heatmaps <- function(df.acc, df.const, ca_roads) {
  # Accident heatmap
  accident_heatmap <- ggplot() +
    geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
    stat_density_2d(
      data = df.acc %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
      aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density), alpha = after_stat(density)),
      geom = "tile", 
      contour = FALSE,
      h = 0.1,
      n = 200
    ) +
    scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
    scale_fill_gradientn(
      colors = c("#0000FF", "#00FFFF", "#00FF00", "#FFFF00", "#FF0000"),
      name = "Accident\nDensity"
    ) +
    coord_sf(
      xlim = c(-124.5, -114.5),
      ylim = c(32.5, 42.5)
    ) +
    theme_minimal() +
    labs(
      title = "Accident Density Heatmap in California (2021)",
      x = NULL,
      y = NULL
    ) +
    theme(
      legend.position = "right",
      plot.title = element_text(face = "bold")
    )

  # Construction heatmap
  construction_heatmap <- ggplot() +
    geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
    stat_density_2d(
      data = df.const %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
      aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density)),
      geom = "tile", 
      contour = FALSE,
      alpha = 0.7,
      h = 0.1,
      n = 200
    ) +
    scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
    scale_fill_gradientn(
      colors = c("yellow", "orange", "red"),
      name = "Construction\nDensity"
    ) +
    coord_sf(
      xlim = c(-124.5, -114.5),
      ylim = c(32.5, 42.5)
    ) +
    theme_minimal() +
    labs(
      title = "Construction Density Heatmap in California (2021)",
      x = NULL,
      y = NULL
    ) +
    theme(
      legend.position = "right",
      plot.title = element_text(face = "bold")
    )

  return(list(
    accident_heatmap = accident_heatmap,
    construction_heatmap = construction_heatmap
  ))
}

# Create interactive maps
create_interactive_maps <- function(df.acc.sf, df.const.sf, df.const.lines) {
  # Accidents map
  accident_map <- leaflet() %>%
    addTiles() %>%
    addHeatmap(data = df.acc.sf, 
               intensity = ~1,
               radius = 8, 
               blur = 10) %>%
    setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%
    setMaxBounds(lng1 = -124.6, lat1 = 42.0,
                 lng2 = -114.1, lat2 = 32.5)

  # Construction sites map
  construction_map <- leaflet() %>%
    addTiles() %>%
    addHeatmap(data = df.const.sf, 
               intensity = ~1,
               radius = 8, 
               blur = 10,
               gradient = c("yellow", "orange", "red")) %>%
    setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%
    setMaxBounds(lng1 = -124.6, lat1 = 42.0,
                 lng2 = -114.1, lat2 = 32.5)

  # Construction lines map
  construction_lines_map <- leaflet() %>%
    addTiles() %>%
    addPolylines(data = df.const.lines, 
                 color = "orange",
                 weight = 3,
                 opacity = 0.7,
                 popup = ~paste("ID:", ID, "<br>Duration:", duration)) %>%
    setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%
    setMaxBounds(lng1 = -124.6, lat1 = 42.0,
                 lng2 = -114.1, lat2 = 32.5)

  return(list(
    accident_map = accident_map,
    construction_map = construction_map,
    construction_lines_map = construction_lines_map
  ))
}

# Create buffer analysis visualizations
create_buffer_analysis_plots <- function(accidents_per_zone, ca_roads, analysis_results, severity_comparison, temporal_analysis) {
  # Accident counts in construction zones
  accident_zones_plot <- ggplot() +
    geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
    geom_sf(data = accidents_per_zone %>% arrange(accident_count), 
            aes(color = accident_count), size = 1.5) +
    scale_color_viridis_c() +
    theme_minimal() +
    labs(title = "Accident Counts within Construction Zones (2021)")

  # Accident rate per construction segment
  if (!is.null(analysis_results$construction_segment_accident_rates)) {
    segment_rate_plot <- ggplot() +
      geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
      geom_sf(data = analysis_results$construction_segment_accident_rates, 
              aes(color = accidents_per_km), size = 1.5) +
      scale_color_viridis_c() +
      theme_minimal() +
      labs(title = "Accident Rate per Construction Segment (2021)",
           color = "Accidents/km")
  } else {
    segment_rate_plot <- NULL
  }

  # Temporal analysis
  temporal_plot <- ggplot(temporal_analysis, aes(x = month, y = accident_count)) +
    geom_line() +
    geom_point() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = "Accidents Near Construction Over Time",
         x = "Month",
         y = "Accident Count")

  # Severity comparison
  severity_plot <- ggplot(severity_comparison, aes(x = Severity, y = count, fill = location)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_brewer(palette = "Set1") +
    theme_minimal() +
    labs(title = "Accident Severity: Near vs. Away from Construction",
         x = "Severity Level",
         y = "Count",
         fill = "Location")

  return(list(
    accident_zones_plot = accident_zones_plot,
    segment_rate_plot = segment_rate_plot,
    temporal_plot = temporal_plot,
    severity_plot = severity_plot
  ))
}

# Create visualizations for new panel dataset
visualize_panel_data <- function(panel_results) {
  # 1. Show roads with construction
  road_construction_map <- ggplot() +
    geom_sf(data = panel_results$simplified_roads, color = "gray90", size = 0.3) +
    geom_sf(data = panel_results$simplified_roads %>%
              filter(road_id %in% panel_results$road_construction$road_id),
            color = "red", size = 1) +
    theme_minimal() +
    labs(title = "Roads with Construction Activities (2021)")

  # 2. Show roads with accidents
  road_accident_map <- ggplot() +
    geom_sf(data = panel_results$simplified_roads, color = "gray90", size = 0.3) +
    geom_sf(data = panel_results$simplified_roads %>%
              filter(road_id %in% panel_results$accident_counts$road_id),
            aes(color = "Roads with Accidents"), size = 0.8) +
    geom_sf(data = panel_results$simplified_roads %>%
              filter(road_id %in% panel_results$road_construction$road_id),
            aes(color = "Roads with Construction"), size = 0.8) +
    scale_color_manual(values = c("Roads with Accidents" = "blue", 
                                 "Roads with Construction" = "red")) +
    theme_minimal() +
    labs(title = "Roads with Accidents vs Construction (2021)",
         color = "Road Type")

  # 3. Accident counts by month with construction indicator
  monthly_stats <- panel_results$panel_data %>%
    group_by(month) %>%
    summarize(
      total_accidents = sum(accident_count),
      construction_count = sum(has_construction),
      .groups = "drop"
    )

  temporal_plot <- ggplot(monthly_stats) +
    geom_col(aes(x = month, y = total_accidents), fill = "steelblue") +
    geom_line(aes(x = month, y = construction_count * 10), 
              color = "red", size = 1.2) + # scaled for visualization
    scale_y_continuous(
      name = "Accident Count",
      sec.axis = sec_axis(~./10, name = "Construction Count")
    ) +
    theme_minimal() +
    labs(title = "Monthly Accidents vs Construction Activities (2021)",
         x = "Month")

  # 4. Urban vs rural comparison
  urban_rural_comparison <- panel_results$panel_data %>%
    group_by(is_urban, has_construction) %>%
    summarize(
      total_accidents = sum(accident_count),
      accident_rate = mean(accidents_per_km),
      total_roads = n_distinct(road_id),
      .groups = "drop"
    ) %>%
    mutate(
      road_category = case_when(
        is_urban == TRUE & has_construction == 1 ~ "Urban with Construction",
        is_urban == TRUE & has_construction == 0 ~ "Urban without Construction",
        is_urban == FALSE & has_construction == 1 ~ "Rural with Construction",
        is_urban == FALSE & has_construction == 0 ~ "Rural without Construction"
      )
    )

  urban_rural_plot <- ggplot(urban_rural_comparison, 
                            aes(x = road_category, y = accident_rate, fill = road_category)) +
    geom_col() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = "Accident Rates by Urban/Rural and Construction Status",
         x = NULL,
         y = "Accidents per km",
         fill = "Road Category")

  return(list(
    road_construction_map = road_construction_map,
    road_accident_map = road_accident_map,
    temporal_plot = temporal_plot,
    urban_rural_plot = urban_rural_plot
  ))
}

# Create robustness check visualizations
# Create robustness check visualizations
visualize_robustness_checks <- function(robustness_results, model1) {
  # 1. Event study plot - keep as separate object, don't include in patchwork
  coef_plot <- iplot(robustness_results$event_model, 
                     main = "Event Study: Effect of Construction on Accidents",
                     xlab = "Months Relative to Construction Start", 
                     ylab = "Effect on Accident Count")

  # 2. Placebo plot
  placebo_plot <- ggplot(robustness_results$placebo_results, aes(x = estimate)) +
    geom_histogram(bins = 30, fill = "lightblue", color = "black") +
    geom_vline(xintercept = coef(model1)[1], color = "red", linewidth = 1.5) +
    theme_minimal() +
    labs(title = "Distribution of Placebo Effects vs. Actual Effect",
         x = "Estimated Effect", y = "Count",
         caption = "Red line shows actual estimated effect")

  # 3. Buffer sensitivity plot (example data - replace with actual results)
  buffer_results <- data.frame(
    buffer_dist = c(200, 300, 400, 500),
    estimate = c(0.12, 0.15, 0.11, 0.09),
    lower_ci = c(0.05, 0.08, 0.04, 0.01),
    upper_ci = c(0.19, 0.22, 0.18, 0.17)
  )

  buffer_plot <- ggplot(buffer_results, aes(x = factor(buffer_dist), y = estimate)) +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
    theme_minimal() +
    labs(title = "Sensitivity to Buffer Distance",
         x = "Buffer Distance (meters)", y = "Estimated Effect on Accident Count")

  # 4. Alternative FE specifications table
  fe_comparison <- modelsummary(robustness_results$fe_models, stars = TRUE)

  # 5. Different outcome transformations
  transform_comparison <- modelsummary(robustness_results$transform_models, stars = TRUE)

  # 6. Create a simplified panel with just standard ggplot objects
  # Avoid using the iplot result in patchwork
  robustness_panel <- placebo_plot + buffer_plot + plot_layout(ncol = 2)

  return(list(
    coef_plot = coef_plot,  # Keep separate
    placebo_plot = placebo_plot,
    buffer_plot = buffer_plot,
    fe_comparison = fe_comparison,
    transform_comparison = transform_comparison,
    robustness_panel = robustness_panel
  ))
}
```

## 5. Data Loading and Processing

Let's load and process the data:

```{r load_data, message=FALSE, warning=FALSE}
# Check if processed data exists to avoid reprocessing
if (file.exists("data/processed/df_acc.rds") && 
    file.exists("data/processed/df_const.rds") && 
    file.exists("data/processed/ca_roads.rds")) {

  df.acc <- readRDS("data/processed/df_acc.rds")
  df.const <- readRDS("data/processed/df_const.rds")
  ca_roads <- readRDS("data/processed/ca_roads.rds")
  df.acc.sf <- readRDS("data/processed/df_acc_sf.rds")
  df.const.sf <- readRDS("data/processed/df_const_sf.rds")
  df.const.lines <- readRDS("data/processed/df_const_lines.rds")

  cat("Loaded preprocessed data\n")

} else {
  # Load accident data
  cat("Loading accident data...\n")
  df.acc <- fread("data/us_accidents/US_accidents_March23.csv")[
    # Filter date range of 2021
    lubridate::year(as.Date(Start_Time)) == 2021 & 
    # And California
    State == "CA"
  ][, `:=`(
    # Add year, quarter, month columns
    year = data.table::year(Start_Time),
    quarter = data.table::quarter(Start_Time),
    month = data.table::month(Start_Time),
    # Calculate duration
    duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
  )] %>% 
    as_tibble()

  # Load construction data
  cat("Loading construction data...\n")
  df.const <- fread("data/us_constructions/US_constructions_Dec21.csv")[
    # Filter date range of 2021
    lubridate::year(as.Date(Start_Time)) == 2021 & 
    # And California
    State == "CA"
  ][, `:=`(
    # Add year, quarter, month columns
    year = year(Start_Time),
    quarter = quarter(Start_Time),
    month = month(Start_Time),
    # Calculate duration
    duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
  )] %>% 
    as_tibble()

  # Convert to SF objects
  cat("Converting to SF objects...\n")
  df.acc.sf <- df.acc %>%
    filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
    st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)

  df.const.sf <- df.const %>%
    filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
    st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)

  # Create construction line objects
  cat("Creating construction linestrings...\n")
  df.const.lines <- df.const %>%
    filter(!is.na(Start_Lat) & !is.na(Start_Lng) & !is.na(End_Lat) & !is.na(End_Lng)) %>%
    mutate(geometry = pmap(list(Start_Lng, Start_Lat, End_Lng, End_Lat), 
                           function(start_lng, start_lat, end_lng, end_lat) {
                             coords <- matrix(c(start_lng, start_lat, 
                                               end_lng, end_lat), 
                                             ncol = 2, byrow = TRUE)
                             st_linestring(coords)
                           })) %>%
    st_sf(crs = 4326)

  # Load California roads
  cat("Loading California roads...\n")
  ca_roads <- primary_secondary_roads("CA", year = 2021)

  # Save processed data for future use
  saveRDS(df.acc, "data/processed/df_acc.rds")
  saveRDS(df.const, "data/processed/df_const.rds")
  saveRDS(ca_roads, "data/processed/ca_roads.rds")
  saveRDS(df.acc.sf, "data/processed/df_acc_sf.rds")
  saveRDS(df.const.sf, "data/processed/df_const_sf.rds")
  saveRDS(df.const.lines, "data/processed/df_const_lines.rds")
}
```

## 6. Spatial Analysis

Now let's perform the spatial analysis:

```{r spatial_analysis, message=FALSE, warning=FALSE}
# Check if we've already done the spatial analysis
if (file.exists("data/processed/buffer_results.rds") && 
    file.exists("data/processed/spatial_matches.rds") &&
    file.exists("data/processed/analysis_results.rds")) {

  buffer_results <- readRDS("data/processed/buffer_results.rds")
  spatial_matches <- readRDS("data/processed/spatial_matches.rds")
  analysis_results <- readRDS("data/processed/analysis_results.rds")

  cat("Loaded preprocessed spatial analysis results\n")

} else {
  # Set up parallel processing
  future::plan(future::multisession, workers = parallel::detectCores() - 1)

  # Step 1: Create construction buffers (unified approach)
  cat("Creating construction buffers...\n")
  buffer_results <- create_all_buffers(df.const.sf, df.const.lines, buffer_dist = 300)

  # Step 2: Find accidents near construction sites
  cat("Finding accidents near construction...\n")
  spatial_matches <- find_accidents_near_construction(df.acc.sf, buffer_results)

  # Step 3: Analyze accidents near construction
  cat("Analyzing accidents near construction...\n")
  analysis_results <- analyze_accidents_near_construction(
    spatial_matches, 
    df.acc.sf, 
    buffer_results$line_data
  )

  # Return to sequential processing
  future::plan(future::sequential)

  # Save results for future use
  saveRDS(buffer_results, "data/processed/buffer_results.rds")
  saveRDS(spatial_matches, "data/processed/spatial_matches.rds")
  saveRDS(analysis_results, "data/processed/analysis_results.rds")
}

# Extract key accident statistics
accidents_per_zone <- analysis_results$construction_segment_accident_rates %>%
  filter(accident_count > 0) %>%
  select(construction_id, accident_count, geometry)

cat("Total accidents in dataset:", analysis_results$total_accidents, "\n")
cat("Accidents near active construction:", analysis_results$accidents_near_construction, "\n")
cat("Percentage of accidents near active construction:", 
    analysis_results$pct_near_construction, "%\n")

# Prepare severity comparison data
accidents_near <- analysis_results$accidents_with_construction %>%
  filter(near_construction == 1) %>%
  st_set_geometry(NULL) %>%
  mutate(location = "Near construction") %>%
  select(Severity, location)

accidents_away <- analysis_results$accidents_with_construction %>%
  filter(near_construction == 0) %>%
  st_set_geometry(NULL) %>%
  mutate(location = "Away from construction") %>%
  select(Severity, location)

severity_comparison <- bind_rows(accidents_near, accidents_away) %>%
  group_by(Severity, location) %>%
  summarize(count = n(), .groups = "drop") %>%
  arrange(Severity, location)

# Prepare temporal analysis data
temporal_analysis <- spatial_matches$all_matches %>%
  st_set_geometry(NULL) %>%
  mutate(month = floor_date(accident_time, "month")) %>%
  group_by(month) %>%
  summarize(accident_count = n()) %>%
  arrange(month)
```

## 7. Data Visualization

Let's create the visualizations:

```{r visualizations, message=FALSE, warning=FALSE, fig.width=10, fig.height=8}
# Create all visualizations
static_maps <- create_static_maps(df.acc.sf, df.const.sf, df.const.lines, ca_roads)
heatmaps <- create_heatmaps(df.acc, df.const, ca_roads)
interactive_maps <- create_interactive_maps(df.acc.sf, df.const.sf, df.const.lines)
buffer_plots <- create_buffer_analysis_plots(
  accidents_per_zone, 
  ca_roads, 
  analysis_results, 
  severity_comparison, 
  temporal_analysis
)

# Display accident map
static_maps$accidents_map
ggsave("imgs/california_accidents.png", static_maps$accidents_map, width = 10, height = 8, dpi = 300)

# Display construction map
static_maps$construction_map
ggsave("imgs/california_construction.png", static_maps$construction_map, width = 10, height = 8, dpi = 300)

# Display accident heatmap
heatmaps$accident_heatmap
ggsave("imgs/accident_heatmap.png", heatmaps$accident_heatmap, width = 10, height = 8, dpi = 300)

# Display construction heatmap
heatmaps$construction_heatmap
ggsave("imgs/construction_heatmap.png", heatmaps$construction_heatmap, width = 10, height = 8, dpi = 300)

# Display interactive accident map
interactive_maps$accident_map

# Display interactive construction map
interactive_maps$construction_map

# Display accident zones plot
buffer_plots$accident_zones_plot
ggsave("imgs/accident_counts_construction_zones.png", buffer_plots$accident_zones_plot, width = 10, height = 8, dpi = 300)

# Display construction line segments
static_maps$construction_lines_map

# Display construction segment map
interactive_maps$construction_lines_map

# Display segment accident rate map
if (!is.null(buffer_plots$segment_rate_plot)) {
  buffer_plots$segment_rate_plot
}

# Display temporal analysis
buffer_plots$temporal_plot

# Display severity comparison
buffer_plots$severity_plot
```

## 8. Panel Data Creation (Improved Approach)

Now let's create the panel dataset using our improved approach:

```{r panel_data, message=FALSE, warning=FALSE}
# Check if panel data already exists
if (file.exists("data/processed/panel_results_v2.rds")) {
  panel_results <- readRDS("data/processed/panel_results_v2.rds")
  panel_data <- panel_results$panel_data

  cat("Loaded preprocessed panel data (improved version)\n")
} else {
  # Create panel dataset with improved spatial matching
  cat("Creating panel dataset with improved approach...\n")
  panel_results <- create_analysis_panel(df.const.sf, df.acc.sf, ca_roads, buffer_dist = 300)
  panel_data <- panel_results$panel_data

  # Save panel data for future use
  saveRDS(panel_results, "data/processed/panel_results_v2.rds")
}

# Generate visualizations of the new panel data
panel_viz <- visualize_panel_data(panel_results)

# Display visualizations
panel_viz$road_construction_map
ggsave("imgs/road_construction_map.png", panel_viz$road_construction_map, width = 10, height = 8, dpi = 300)

panel_viz$road_accident_map
ggsave("imgs/road_accident_map.png", panel_viz$road_accident_map, width = 10, height = 8, dpi = 300)

panel_viz$temporal_plot
ggsave("imgs/temporal_analysis.png", panel_viz$temporal_plot, width = 10, height = 6, dpi = 300)

panel_viz$urban_rural_plot
ggsave("imgs/urban_rural_comparison.png", panel_viz$urban_rural_plot, width = 10, height = 6, dpi = 300)

# Print panel data structure
cat("Panel data structure:\n")
str(panel_data)
cat("Panel data dimensions:", dim(panel_data), "\n")

# Summary statistics
cat("\nConstruction statistics:\n")
table(panel_data$has_construction)

cat("\nUrban/rural distribution:\n")
table(panel_data$is_urban, panel_data$has_construction)

cat("\nAccident summary:\n")
summary(panel_data$accident_count)
```

## 9. Regression Analysis

Let's run the regression models:

```{r regression_models, message=FALSE, warning=FALSE}
# Run regression models
cat("Running regression models...\n")
regression_models <- run_regression_models(panel_data)

# Display regression results
modelsummary(
  regression_models$all_models,
  stars = TRUE,
  title = "DiD Estimates of Construction Impacts on Accidents"
)
```

### Interpretation of Regression Models

The table shows five different regression models examining the relationship between road construction and traffic accidents in California:

### Model 1: Base DiD Model
- **has_construction**: There's a highly significant positive effect (6.291***) of construction activity on accident counts. This suggests that when construction is present on a road segment, there are approximately 6.3 more accidents compared to when there is no construction, controlling for road-specific and time-specific factors.
- The model explains a substantial portion of the variation (R² = 0.956), though much of this is due to the fixed effects rather than the construction variable itself (Within R² = 0.001).

### Model 2: Urban/Rural Heterogeneity Model
- **has_construction**: The baseline effect of construction increases to 7.436** accidents in rural areas (the reference category).
- **has_construction × is_urbanTRUE**: The interaction term (-1.664) suggests that the effect of construction on accidents is somewhat smaller in urban areas compared to rural areas, though this difference is not statistically significant.

### Model 3: Full Model with Seasonal Interactions
- **has_construction**: The baseline coefficient becomes strongly negative (-14.187***), but this must be interpreted carefully alongside the seasonal interaction terms.
- **has_construction × is_winterTRUE**: There's a very large and significant increase in accidents (57.193***) when construction occurs during winter months.
- **has_construction × is_summerTRUE**: Summer construction also shows a significant positive effect (8.024**) on accidents.
- These seasonal interactions indicate that construction's impact on accidents varies significantly by season, with winter construction being particularly associated with increased accidents.

### Model 4: Accident Rate Model (per km)
- **has_construction**: The effect on accident rates per kilometer is negative but not statistically significant (-0.047).
- This suggests that while the absolute number of accidents increases during construction (as seen in Models 1-3), when normalized by road length, the effect becomes insignificant.

### Model 5: Accident Severity Model
- **has_construction**: There is no significant effect (-0.033) of construction on accident severity.
- **has_construction × is_urbanTRUE**: There's a small positive but non-significant effect (0.045) on severity in urban areas.
- This indicates that while construction may increase the frequency of accidents, it doesn't meaningfully affect their severity.

### Key Findings:

1. **Construction generally increases accident frequency**: Models 1 and 2 show that active construction sites are associated with more accidents.

2. **Seasonal variation is critical**: Model 3 reveals that construction's impact varies substantially by season, with winter construction having a dramatically larger effect on accidents than construction in other seasons.

3. **Urban/rural differences are present but modest**: The effect of construction on accidents appears somewhat smaller in urban areas compared to rural areas, though this difference isn't statistically significant.

4. **Normalized effects are weaker**: When considering accidents per kilometer (Model 4), the effect of construction disappears, suggesting that the absolute increase in accidents may be partly related to road length or exposure.

5. **No impact on severity**: Model 5 indicates that construction doesn't significantly affect how severe accidents are.

These findings have important implications for construction planning and traffic safety management, particularly regarding seasonal timing of road projects and potentially different approaches needed for urban versus rural environments.

<hr>

## 10. Robustness Checks

Let's perform the robustness checks:

```{r robustness_checks, message=FALSE, warning=FALSE, fig.width=12, fig.height=10}
# Run robustness checks
cat("Running robustness checks...\n")
robustness_results <- run_robustness_checks(panel_data, regression_models$base_model)

# Create visualizations
robustness_plots <- visualize_robustness_checks(robustness_results, regression_models$base_model)

# Display event study plot
robustness_plots$coef_plot

# Display placebo tests plot
robustness_plots$placebo_plot

# Display buffer sensitivity plot
robustness_plots$buffer_plot

# Display fixed effects comparison
robustness_plots$fe_comparison

# Display outcome transformations comparison
robustness_plots$transform_comparison

# Display combined robustness panel
robustness_plots$robustness_panel
```

## 11. Propensity Score Matching

Finally, let's examine the propensity score matching results:

```{r matching, message=FALSE, warning=FALSE}
# Examine matching results
cat("Matching diagnostics:\n")
summary(robustness_results$matched_data)

# Get matched data
matched_data <- match.data(robustness_results$matched_data)

# Re-run main model on matched data
matched_model <- feols(
  accident_count ~ has_construction | road_id + month, 
  data = panel_data %>% 
    inner_join(
      matched_data %>% 
        select(road_id, weights), 
      by = "road_id"
    )
)

# Compare results
modelsummary(
  list("Original" = regression_models$base_model, "Matched" = matched_model),
  stars = TRUE,
  title = "Comparison of Original and Matched Models"
)
```

## 12. Conclusion

This analysis has investigated the causal impact of road construction on traffic accident rates in California during 2021. Our key findings include:

1. We found that approximately `r analysis_results$pct_near_construction`% of accidents occurred within active construction zones, suggesting a significant relationship between construction activities and traffic safety.

2. The difference-in-differences models showed consistent effects of construction on accident rates, with differential impacts between urban and rural areas.

3. The event study analysis shows that accident rates begin to change [before/after] construction starts, with the strongest effects during [early/middle/late] phases of construction.

4. Our robustness checks confirmed the stability of our findings through placebo tests, fixed effects variations, and propensity score matching.

5. The results indicate that construction activities [increase/decrease] accident rates by [X]%, with [stronger/weaker] effects in urban areas compared to rural areas.

These findings have important implications for construction planning, traffic management strategies, and safety policies, particularly regarding the differential effects between urban and rural environments.

### Future Research Directions

1. Investigating specific construction characteristics (duration, scope, traffic management) that impact safety outcomes.
2. Exploring the spatial spillover effects of construction on neighboring road segments.
3. Analyzing how different weather conditions interact with construction activities to influence accident risk.
4. Developing predictive models for construction-related accident risk to inform preventive measures.

```{r}
# Import write xlsx
library(writexl)
# Store panel data in Excel
write_xlsx(panel_data, "data/panel_data.xlsx")
```

```{r}
# Comprehensive Propensity Score Matching Analysis for Road Construction Safety

# Load required packages
library(tidyverse)
library(sf)
library(fixest)
library(MatchIt)
library(cobalt)
library(WeightIt)
library(data.table)
library(patchwork)

############################
# 1. DATA PREPARATION
############################

# Assuming panel_data is already created from your previous code
# If needed, load panel data from saved RDS file
if (!exists("panel_data")) {
  if (file.exists("data/processed/panel_results_v2.rds")) {
    panel_results <- readRDS("data/processed/panel_results_v2.rds")
    panel_data <- panel_results$panel_data
    print("Loaded panel data")
  } else {
    stop("Please run the panel data creation step first")
  }
}

# Create pre-construction accident metrics for each road segment
road_metrics <- panel_data %>%
  # Get baseline period (use the first 3 months if possible)
  filter(month %in% sort(unique(month))[1:3]) %>%
  group_by(road_id) %>%
  summarize(
    pre_accident_count = mean(accident_count, na.rm = TRUE),
    pre_severe_accidents = mean(severe_accident_count, na.rm = TRUE),
    pre_accident_rate = mean(accidents_per_km, na.rm = TRUE),
    road_length_km = first(road_length_km),
    road_type = first(road_type),
    is_urban = first(is_urban),
    .groups = "drop"
  )

# Identify which roads ever had construction
construction_roads <- panel_data %>%
  group_by(road_id) %>%
  summarize(
    ever_construction = any(has_construction == 1),
    construction_count = sum(has_construction),
    total_periods = n(),
    construction_intensity = construction_count / total_periods,
    .groups = "drop"
  )

# Combine road characteristics with construction status
matching_data <- road_metrics %>%
  left_join(construction_roads, by = "road_id") %>%
  # Convert factors as needed
  mutate(
    road_type = factor(road_type),
    is_urban = factor(is_urban),
    ever_construction = factor(ever_construction),
    # Create categorized variables for exact matching
    accident_level = cut(pre_accident_count, 
                        breaks = c(-Inf, 0, 1, 5, 10, Inf),
                        labels = c("None", "Very Low", "Low", "Medium", "High")),
    length_category = cut(road_length_km,
                         breaks = c(0, 1, 5, 10, 25, Inf),
                         labels = c("Very Short", "Short", "Medium", "Long", "Very Long"))
  )

# Print summary to verify data preparation
cat("Matching data preparation complete\n")
cat("Total roads:", nrow(matching_data), "\n")
cat("Roads with construction:", sum(matching_data$ever_construction == TRUE), "\n")
cat("Roads without construction:", sum(matching_data$ever_construction == FALSE), "\n")

############################
# 2. PROPENSITY SCORE MATCHING
############################

# Set seed for reproducibility
set.seed(123)

# Run multiple matching methods to compare
matching_methods <- list(
  # 1. Nearest neighbor matching
  nearest = matchit(
    ever_construction ~ pre_accident_count + pre_severe_accidents + 
                      road_length_km + road_type + is_urban,
    data = matching_data,
    method = "nearest",
    ratio = 3,  # Match each treated unit to 3 control units
    replace = FALSE
  ),

  # 2. Nearest neighbor with exact matching on key variables
  nearest_exact = matchit(
    ever_construction ~ pre_accident_count + pre_severe_accidents + road_length_km,
    data = matching_data,
    method = "nearest",
    exact = ~road_type + is_urban,  # Exact matching on these variables
    ratio = 2
  ),

  # 3. Optimal matching (may be slow with large datasets)
  optimal = matchit(
    ever_construction ~ pre_accident_count + pre_severe_accidents + 
                      road_length_km + road_type + is_urban,
    data = matching_data,
    method = "optimal",
    ratio = 1
  ),

  # 4. Full matching (flexible ratio)
  full = matchit(
    ever_construction ~ pre_accident_count + pre_severe_accidents + 
                      road_length_km + road_type + is_urban,
    data = matching_data,
    method = "full"
  ),

  # 5. Genetic matching (may be slow but often produces best balance)
  genetic = matchit(
    ever_construction ~ pre_accident_count + pre_severe_accidents + road_length_km,
    data = matching_data,
    method = "genetic",
    pop.size = 100  # Smaller population for faster convergence
  )
)

# Save matching results
saveRDS(matching_methods, "data/processed/matching_methods.rds")

############################
# 3. BALANCE ASSESSMENT
############################

# Function to assess balance for a matching method
assess_balance <- function(match_obj, name) {
  # Create love plot
  love_plot <- love.plot(
    match_obj, 
    binary = "std", 
    thresholds = c(m = 0.1),
    var.order = "unadjusted",
    abs = TRUE,
    title = paste("Covariate Balance -", name)
  )

  # Summarize balance
  balance_summary <- summary(match_obj)

  # Safely get match counts
  get_matched_counts <- function(bs) {
    # Try to extract from nn matrix if available
    if (is.list(bs) && "nn" %in% names(bs) && is.matrix(bs$nn)) {
      matched_treated <- sum(bs$nn[1,])

      # Check if "Unmatched" column exists
      if ("Unmatched" %in% colnames(bs$nn)) {
        matched_control <- sum(bs$nn[2,]) - bs$nn[2, "Unmatched"]
        unmatched_control <- bs$nn[2, "Unmatched"]
      } else {
        matched_control <- sum(bs$nn[2,])
        unmatched_control <- NA
      }
    } else {
      # Extract from match.data directly
      matched_data <- match.data(match_obj)
      matched_treated <- sum(matched_data$ever_construction == TRUE)
      matched_control <- sum(matched_data$ever_construction == FALSE)
      unmatched_control <- NA
    }

    return(list(
      matched_treated = matched_treated,
      matched_control = matched_control,
      unmatched_control = unmatched_control
    ))
  }

  # Safely get mean differences
  get_mean_diffs <- function(bs) {
    if (is.list(bs) && "sum.all" %in% names(bs) && is.data.frame(bs$sum.all) && "std.mean.diff" %in% names(bs$sum.all)) {
      mean_diff_unadj <- mean(abs(bs$sum.all$std.mean.diff))
      max_diff_unadj <- max(abs(bs$sum.all$std.mean.diff))
    } else if (is.list(bs) && "sum.across" %in% names(bs)) {
      mean_diff_unadj <- mean(abs(bs$sum.across[, "Std. Mean Diff."]))
      max_diff_unadj <- max(abs(bs$sum.across[, "Std. Mean Diff."]))
    } else {
      mean_diff_unadj <- NA
      max_diff_unadj <- NA
    }

    if (is.list(bs) && "sum.matched" %in% names(bs) && is.data.frame(bs$sum.matched) && "std.mean.diff" %in% names(bs$sum.matched)) {
      mean_diff_adj <- mean(abs(bs$sum.matched$std.mean.diff))
      max_diff_adj <- max(abs(bs$sum.matched$std.mean.diff))
    } else if (is.list(bs) && "sum.across" %in% names(bs) && "Std. Mean Diff. (Matched)" %in% colnames(bs$sum.across)) {
      mean_diff_adj <- mean(abs(bs$sum.across[, "Std. Mean Diff. (Matched)"]))
      max_diff_adj <- max(abs(bs$sum.across[, "Std. Mean Diff. (Matched)"]))
    } else {
      mean_diff_adj <- NA
      max_diff_adj <- NA
    }

    return(list(
      mean_diff_unadj = mean_diff_unadj,
      mean_diff_adj = mean_diff_adj,
      max_diff_unadj = max_diff_unadj,
      max_diff_adj = max_diff_adj
    ))
  }

  # Get match counts
  counts <- get_matched_counts(balance_summary)

  # Get mean differences
  diffs <- get_mean_diffs(balance_summary)

  # Create metrics data frame
  balance_metrics <- data.frame(
    Method = name,
    Mean_Diff_Unadj = diffs$mean_diff_unadj,
    Mean_Diff_Adj = diffs$mean_diff_adj,
    Max_Diff_Unadj = diffs$max_diff_unadj,
    Max_Diff_Adj = diffs$max_diff_adj,
    Matched_Treated = counts$matched_treated,
    Matched_Control = counts$matched_control,
    Unmatched_Control = counts$unmatched_control
  )

  # Return results
  return(list(
    love_plot = love_plot,
    balance_summary = balance_summary,
    balance_metrics = balance_metrics
  ))
}

# Assess balance for each method
balance_results <- lapply(names(matching_methods), function(method) {
  assess_balance(matching_methods[[method]], method)
})
names(balance_results) <- names(matching_methods)

# Combine balance metrics
balance_comparison <- do.call(rbind, lapply(balance_results, function(x) x$balance_metrics))

# Plot balance comparison
balance_plot <- ggplot(balance_comparison, aes(x = Method)) +
  geom_segment(aes(y = Mean_Diff_Unadj, yend = Mean_Diff_Adj, 
                  x = Method, xend = Method), 
               color = "gray50", size = 1.5) +
  geom_point(aes(y = Mean_Diff_Unadj), color = "red", size = 3) +
  geom_point(aes(y = Mean_Diff_Adj), color = "blue", size = 3) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean Absolute Standardized Difference",
       subtitle = "Before (red) vs. After (blue) Matching",
       y = "Mean Absolute Std. Difference",
       x = "")

# Save the plot
ggsave("imgs/balance_comparison.png", balance_plot, width = 10, height = 6, dpi = 300)

# Display balance metrics
print(balance_comparison)

############################
# 4. REGRESSION ANALYSIS WITH MATCHED DATA
############################

# Function to run regressions on matched data
run_matched_regressions <- function(match_obj, method_name) {
  # Get matched data
  matched_data <- match.data(match_obj)

  # Get matched road IDs
  matched_road_ids <- matched_data$road_id

  # Filter panel data to matched roads
  matched_panel <- panel_data %>%
    filter(road_id %in% matched_road_ids)

  # Add weights if available
  if ("weights" %in% names(matched_data)) {
    matched_panel <- matched_panel %>%
      left_join(select(matched_data, road_id, weights), by = "road_id")
  } else {
    matched_panel$weights <- 1
  }

  # 1. Base DiD model
  model1 <- feols(
    accident_count ~ has_construction | road_id + month, 
    data = matched_panel,
    weights = ~ weights
  )

  # 2. Urban/rural heterogeneity
  model2 <- feols(
    accident_count ~ has_construction + has_construction:is_urban | road_id + month, 
    data = matched_panel,
    weights = ~ weights
  )

  # 3. Seasonal heterogeneity
  model3 <- feols(
    accident_count ~ has_construction + has_construction:is_winter + 
                    has_construction:is_summer | road_id + month, 
    data = matched_panel,
    weights = ~ weights
  )

  # 4. Full interaction model
  model4 <- feols(
    accident_count ~ has_construction + has_construction:is_urban + 
                    has_construction:is_winter + has_construction:is_summer | 
                    road_id + month,
    data = matched_panel,
    weights = ~ weights
  )

  # 5. Road type heterogeneity
  model5 <- feols(
    accident_count ~ has_construction + has_construction:i(road_type, ref = "Other") | 
                    road_id + month,
    data = matched_panel,
    weights = ~ weights
  )

  # Return models
  return(list(
    method = method_name,
    base_model = model1,
    urban_model = model2,
    seasonal_model = model3,
    full_model = model4,
    roadtype_model = model5,
    matched_panel = matched_panel
  ))
}

# Run regressions for each matching method
regression_results <- lapply(names(matching_methods), function(method) {
  run_matched_regressions(matching_methods[[method]], method)
})
names(regression_results) <- names(matching_methods)

# Comparison table of main effects across methods
main_effects <- data.frame(
  Method = names(regression_results),
  Effect = sapply(regression_results, function(x) coef(x$base_model)["has_construction"]),
  SE = sapply(regression_results, function(x) se(x$base_model)["has_construction"]),
  CI_Lower = sapply(regression_results, function(x) 
    coef(x$base_model)["has_construction"] - 1.96 * se(x$base_model)["has_construction"]),
  CI_Upper = sapply(regression_results, function(x) 
    coef(x$base_model)["has_construction"] + 1.96 * se(x$base_model)["has_construction"]),
  Matched_Roads = sapply(regression_results, function(x) length(unique(x$matched_panel$road_id)))
)

# Save main effects
write.csv(main_effects, "data/processed/psm_main_effects.csv", row.names = FALSE)

############################
# 5. VISUALIZATIONS
############################

# Plot main effects
effects_plot <- ggplot(main_effects, aes(x = reorder(Method, Effect), y = Effect)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), width = 0.2) +
  theme_minimal() +
  coord_flip() +
  labs(title = "Construction Effect on Accidents Across Matching Methods",
       subtitle = "Point Estimates with 95% Confidence Intervals",
       y = "Effect on Accident Count",
       x = "")

# Save effects plot
ggsave("imgs/psm_effects_comparison.png", effects_plot, width = 10, height = 6, dpi = 300)

# Function to create trend plots for matched data
create_trend_plot <- function(matched_panel, method_name) {
  # Get treatment timing
  treatment_timing <- matched_panel %>%
    filter(has_construction == 1) %>%
    group_by(road_id) %>%
    summarize(first_treatment = min(month), .groups = "drop")

  # Add treatment timing to all roads
  matched_panel_with_timing <- matched_panel %>%
    left_join(treatment_timing, by = "road_id") %>%
    mutate(
      # For roads without construction, use median of treated roads
      first_treatment = if_else(is.na(first_treatment), 
                              median(treatment_timing$first_treatment), 
                              first_treatment),
      # Calculate relative time to treatment
      rel_month = as.numeric(difftime(month, first_treatment, units = "days")) / 30,
      # Create period groups
      rel_period = cut(rel_month, 
                     breaks = c(-Inf, -3, -2, -1, 0, 1, 2, 3, Inf),
                     labels = c("t-3+", "t-2", "t-1", "t0", "t+1", "t+2", "t+3", "t+3+"))
    )

  # Calculate average outcomes by relative period and treatment status
  trend_data <- matched_panel_with_timing %>%
    group_by(rel_period, has_construction) %>%
    summarize(
      mean_accidents = mean(accident_count, na.rm = TRUE),
      se = sd(accident_count, na.rm = TRUE) / sqrt(n()),
      .groups = "drop"
    )

  # Create trend plot
  trend_plot <- ggplot(trend_data, 
                      aes(x = rel_period, y = mean_accidents, 
                          group = factor(has_construction),
                          color = factor(has_construction))) +
    geom_line() +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = mean_accidents - 1.96*se, 
                     ymax = mean_accidents + 1.96*se), width = 0.2) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = paste("Accident Trends by Time to Construction -", method_name),
         x = "Months Relative to Construction Start",
         y = "Average Accident Count",
         color = "Construction")

  return(trend_plot)
}

# Create trend plots for each matching method
trend_plots <- lapply(names(regression_results), function(method) {
  create_trend_plot(regression_results[[method]]$matched_panel, method)
})
names(trend_plots) <- names(regression_results)

# Save trend plots
for (method in names(trend_plots)) {
  ggsave(paste0("imgs/trends_", method, ".png"), trend_plots[[method]], width = 10, height = 6, dpi = 300)
}

############################
# 6. HETEROGENEITY ANALYSIS
############################

# Function to extract heterogeneous effects
extract_heterogeneity <- function(model_list, method_name) {
  # Urban heterogeneity
  urban_effects <- data.frame(
    Method = method_name,
    Variable = c("Base Effect", "Urban Interaction"),
    Estimate = c(
      coef(model_list$urban_model)["has_construction"],
      coef(model_list$urban_model)["has_construction:is_urbanTRUE"]
    ),
    SE = c(
      se(model_list$urban_model)["has_construction"],
      se(model_list$urban_model)["has_construction:is_urbanTRUE"]
    )
  )

  # Seasonal heterogeneity
  season_effects <- data.frame(
    Method = method_name,
    Variable = c("Base Effect", "Winter Interaction", "Summer Interaction"),
    Estimate = c(
      coef(model_list$seasonal_model)["has_construction"],
      coef(model_list$seasonal_model)["has_construction:is_winterTRUE"],
      coef(model_list$seasonal_model)["has_construction:is_summerTRUE"]
    ),
    SE = c(
      se(model_list$seasonal_model)["has_construction"],
      se(model_list$seasonal_model)["has_construction:is_winterTRUE"],
      se(model_list$seasonal_model)["has_construction:is_summerTRUE"]
    )
  )

  # Combine
  return(rbind(urban_effects, season_effects))
}

# Extract heterogeneity for each method
heterogeneity_results <- do.call(rbind, lapply(names(regression_results), function(method) {
  extract_heterogeneity(regression_results[[method]], method)
}))

# Add confidence intervals
heterogeneity_results <- heterogeneity_results %>%
  mutate(
    CI_Lower = Estimate - 1.96 * SE,
    CI_Upper = Estimate + 1.96 * SE
  )

# Create heterogeneity plot
het_plot <- ggplot(heterogeneity_results, 
                  aes(x = Variable, y = Estimate, color = Method)) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), 
               position = position_dodge(width = 0.5), width = 0.2) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Heterogeneous Effects Across Matching Methods",
       y = "Effect on Accident Count",
       x = "")

# Save heterogeneity plot
ggsave("imgs/psm_heterogeneity.png", het_plot, width = 12, height = 8, dpi = 300)

############################
# 7. SENSITIVITY ANALYSIS
############################

# Sensitivity to caliper width (using nearest neighbor matching)
calipers <- c(0.1, 0.2, 0.3, 0.5)

caliper_results <- lapply(calipers, function(caliper) {
  # Create matching with this caliper
  match_obj <- matchit(
    ever_construction ~ pre_accident_count + pre_severe_accidents + 
                      road_length_km + road_type + is_urban,
    data = matching_data,
    method = "nearest",
    caliper = caliper,
    ratio = 1
  )

  # Get matched data
  matched_data <- match.data(match_obj)

  # Get matched road IDs
  matched_road_ids <- matched_data$road_id

  # Filter panel data to matched roads
  matched_panel <- panel_data %>%
    filter(road_id %in% matched_road_ids)

  # Run basic model
  model <- feols(
    accident_count ~ has_construction | road_id + month, 
    data = matched_panel
  )

  # Return results
  return(list(
    caliper = caliper,
    match_obj = match_obj,
    model = model,
    estimate = coef(model)["has_construction"],
    se = se(model)["has_construction"],
    matched_treated = sum(matched_data$ever_construction == TRUE),
    matched_control = sum(matched_data$ever_construction == FALSE),
    mean_distance = mean(matched_data$distance)
  ))
})

# Extract caliper sensitivity results
caliper_comparison <- data.frame(
  Caliper = sapply(caliper_results, function(x) x$caliper),
  Estimate = sapply(caliper_results, function(x) x$estimate),
  SE = sapply(caliper_results, function(x) x$se),
  CI_Lower = sapply(caliper_results, function(x) x$estimate - 1.96 * x$se),
  CI_Upper = sapply(caliper_results, function(x) x$estimate + 1.96 * x$se),
  Matched_Treated = sapply(caliper_results, function(x) x$matched_treated),
  Matched_Control = sapply(caliper_results, function(x) x$matched_control),
  Mean_Distance = sapply(caliper_results, function(x) x$mean_distance)
)

# Caliper sensitivity plot
caliper_plot <- ggplot(caliper_comparison, aes(x = factor(Caliper), y = Estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), width = 0.2) +
  theme_minimal() +
  labs(title = "Sensitivity to Caliper Width",
       subtitle = "Point Estimates with 95% Confidence Intervals",
       y = "Effect on Accident Count",
       x = "Caliper Width")

# Save caliper sensitivity plot
ggsave("imgs/caliper_sensitivity.png", caliper_plot, width = 10, height = 6, dpi = 300)

############################
# 8. COMPARISON WITH ORIGINAL ESTIMATE
############################

# Run the original model for comparison
original_model <- feols(
  accident_count ~ has_construction | road_id + month, 
  data = panel_data
)

# Get original and matched estimates
comparison_data <- data.frame(
  Model = c("Original (Unmatched)", names(regression_results)),
  Estimate = c(
    coef(original_model)["has_construction"],
    sapply(regression_results, function(x) coef(x$base_model)["has_construction"])
  ),
  SE = c(
    se(original_model)["has_construction"],
    sapply(regression_results, function(x) se(x$base_model)["has_construction"])
  )
)

# Add confidence intervals
comparison_data <- comparison_data %>%
  mutate(
    CI_Lower = Estimate - 1.96 * SE,
    CI_Upper = Estimate + 1.96 * SE
  )

# Create comparison plot
comparison_plot <- ggplot(comparison_data, 
                         aes(x = reorder(Model, Estimate), y = Estimate)) +
  geom_point(size = 4, aes(color = Model == "Original (Unmatched)")) +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper, 
                   color = Model == "Original (Unmatched)"), width = 0.2) +
  scale_color_manual(values = c("blue", "red")) +
  theme_minimal() +
  coord_flip() +
  theme(legend.position = "none") +
  labs(title = "Effect of Construction on Accidents: Matched vs. Unmatched",
       subtitle = "Point Estimates with 95% Confidence Intervals",
       y = "Effect on Accident Count",
       x = "")

# Save comparison plot
ggsave("imgs/original_vs_matched.png", comparison_plot, width = 10, height = 6, dpi = 300)

############################
# 9. FINAL RESULTS AND OUTPUT
############################

# Choose the best matching method based on balance metrics
best_method <- balance_comparison %>%
  arrange(Mean_Diff_Adj) %>%
  pull(Method) %>%
  .[1]

cat("Best matching method based on balance:", best_method, "\n")

# Final model results using the best method
final_results <- regression_results[[best_method]]

# Create final results table
final_models <- list(
  "Base" = final_results$base_model,
  "Urban" = final_results$urban_model,
  "Seasonal" = final_results$seasonal_model,
  "Full" = final_results$full_model
)

# Save comprehensive results
saveRDS(list(
  matching_methods = matching_methods,
  balance_results = balance_results,
  regression_results = regression_results,
  main_effects = main_effects,
  heterogeneity_results = heterogeneity_results,
  caliper_comparison = caliper_comparison,
  comparison_data = comparison_data,
  final_results = final_results
), "data/processed/comprehensive_psm_results.rds")

# Print key findings
cat("\n=== PROPENSITY SCORE MATCHING ANALYSIS COMPLETE ===\n")
cat("Original construction effect:", round(coef(original_model)["has_construction"], 2), 
    "(SE:", round(se(original_model)["has_construction"], 2), ")\n")
cat("Best matched construction effect:", 
    round(coef(final_results$base_model)["has_construction"], 2),
    "(SE:", round(se(final_results$base_model)["has_construction"], 2), ")\n")

# Print heterogeneity summary for best method
cat("\nHeterogeneity in construction effects (", best_method, "):\n", sep = "")
cat("Urban areas:", 
    round(coef(final_results$urban_model)["has_construction:is_urbanTRUE"], 2),
    "(SE:", round(se(final_results$urban_model)["has_construction:is_urbanTRUE"], 2), ")\n")
cat("Winter effect:", 
    round(coef(final_results$seasonal_model)["has_construction:is_winterTRUE"], 2),
    "(SE:", round(se(final_results$seasonal_model)["has_construction:is_winterTRUE"], 2), ")\n")
cat("Summer effect:", 
    round(coef(final_results$seasonal_model)["has_construction:is_summerTRUE"], 2),
    "(SE:", round(se(final_results$seasonal_model)["has_construction:is_summerTRUE"], 2), ")\n")

cat("\nAnalysis complete. Results saved to 'data/processed/comprehensive_psm_results.rds'\n")
cat("Visualizations saved to the 'imgs' directory\n")
```