### 2025-03-25 final/finalSIMON_MoritzPeistNoemiLucchiSimonVellin .Rmd
Duplicated .Rmd file to have a fresh start

---
title: "Geospatial Data Science - Final project"
author:
- Moritz Peist
- Noemi Lucchi
- Simon Vellin
subtitle: Analysis of US road accidents and construction sites
output:
  html_document:
    toc: true
    df_print: paged
  html_notebook:
    df_print: paged
    toc: true
---

<hr>

```{css, echo=FALSE}
h1.title {
  text-align: center;
}

h4.author, h4.date {
  text-align: center;
}

.author {
  text-align: center;
}

.date {
  text-align: center;
}

/* Add styling for subtitle */
h3.subtitle {
  text-align: center;
  font-style: italic;
  margin-top: -10px;
  color: #606060;
}
```

<div style="text-align: center;">
  <img src="../Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
</div>

<hr>

Load packages

```{r Packages}
library(data.table)   # For faster data manipulation
library(tidyverse)    # For data manipulation and visualization
library(sf)           # For spatial data handling
library(leaflet)      # For interactive maps
library(leaflet.extras) # For additional leaflet features
library(mapview)      # For easier map visualization
library(tmap)         # For thematic maps
library(tigris)       # For US road networks
library(future)       # For parallel processing
library(future.apply) # For parallel processing with apply functions
library(sf)           # For spatial data handling
library(RPostgres)    # For PostgreSQL connection
library(DBI)          # For database interface

# Create directories if they don't exist
if (!dir.exists("./data/tigris")) {
  dir.create("./data/tigris", recursive = TRUE)
}

# Set custom cache directory (optional)
options(tigris_cache_dir = "./data/tigris")
# Configure tigris to use caching
options(tigris_use_cache = TRUE)
```

# Set-up convenience functions for parallel processing (as data is quite large)
```{r Parallel Buffering}
# Function to create buffers in batches with proper projection
create_buffers_in_batches <- function(sf_object, buffer_dist, batch_size = 500) {
  # First, reproject to an appropriate projected CRS for the region
  # For California, UTM Zone 10N (EPSG:26910) or 11N (EPSG:26911) works well
  # If we're unsure about your region, a web mercator projection (3857) is a reasonable default
  message("Reprojecting data to a meter-based CRS...")
  sf_object_projected <- st_transform(sf_object, 3310)  # Web Mercator

  n_features <- nrow(sf_object_projected)
  n_batches <- ceiling(n_features / batch_size)

  # Create empty list to store results
  buffer_list <- vector("list", n_batches)

  # Process in batches
  for (i in 1:n_batches) {
    start_idx <- (i-1) * batch_size + 1
    end_idx <- min(i * batch_size, n_features)

    cat(sprintf("Processing batch %d of %d (features %d to %d)\n", 
                i, n_batches, start_idx, end_idx))

    # Extract batch
    batch <- sf_object_projected[start_idx:end_idx, ]

    # Create buffer (with parallel processing within sf)
    buffer_list[[i]] <- st_buffer(batch, dist = buffer_dist)
  }

  # Combine results
  result <- do.call(rbind, buffer_list)

  # Reproject back to original CRS if needed
  message("Reprojecting results back to original CRS...")
  st_transform(result, st_crs(sf_object))
}
```

Load data (filter for California for now to limit data set size)

```{r Data Load Accidents}
# Efficient approach
df.acc <- fread("data/us_accidents/US_accidents_March23.csv")[
  # Filter date range of 2021
  lubridate::year(as.Date(Start_Time)) == 2021 & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = data.table::year(Start_Time),
  quarter = data.table::quarter(Start_Time),
  month = data.table::month(Start_Time),
  # Calculate duration (assuming End_Time exists in the dataset)
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()  # Convert to tibble only at the end for performance
```

```{r Data Load Constructions}
df.const <- fread("data/us_constructions/US_constructions_Dec21.csv")[
  # Filter date range of 2021
  lubridate::year(as.Date(Start_Time)) == 2021 & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = year(Start_Time),
  quarter = quarter(Start_Time),
  month = month(Start_Time),
  # Calculate duration (assuming End_Time exists in the dataset)
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()  # Convert to tibble only at the end for performance
```

Convert dataframes to sf objects

```{r Geospatial Data}
# Convert accident data to sf object
df.acc.sf <- df.acc %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)

# Convert construction data to sf object
df.const.sf <- df.const %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326)
```

```{r Data Load Roads}
# Get US roads (this can be slow for the entire US, so we might want to limit by state)
# We also only get Primary and Secondary roads as they are the most important (i.e., where most constructions and accidents happen). Otherwise, the dataset would become to large (1.1GB)

ca_roads <- primary_secondary_roads("CA", year = 2021)

# Below code to get ALL roads for California (not recommended due to size)

# First get all counties in California
#ca_counties <- counties("CA", year = 2021)

# Then get roads for each county
#ca_roads_list <- lapply(ca_counties$COUNTYFP, function(county_code) {
#  roads("CA", county = county_code, year = 2021)
#})

# Optionally combine all county road data into one object
#ca_roads <- do.call(rbind, ca_roads_list)
```

```{r Static Maps}
# Static map for Accidents with year coloring
# First, store your plot in a variable
ca_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.acc.sf %>% filter(State == "CA"), 
          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
  theme_minimal() +
  labs(title = "US Road Accidents by Year (2016-2021)") +
  theme(legend.position = "bottom")
print(ca_plot)
# Then save it using ggsave
ggsave(filename = "imgs/california_accidents.png", 
       plot = ca_plot,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution

# Construction sites map with year coloring
# First, store our plot in a variable
ca_const_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.const.sf %>% filter(State == "CA"), 
          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
  theme_minimal() +
  labs(title = "US Construction Sites by Year (2016-2021)") +
  theme(legend.position = "bottom")
print(ca_const_plot)
# Then save it using ggsave
ggsave(filename = "imgs/california_construction.png", 
       plot = ca_const_plot,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution
```

```{r Heatmap static}
# Create heatmap of accidents using ggplot2
# Modified accident heatmap with transparency for low density areas
accident_heatmap <- ggplot() +
  # Add California roads as base layer
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  # Create density heatmap with transparency threshold
  stat_density_2d(
    data = df.acc %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density), alpha = after_stat(density)),
    geom = "tile", 
    contour = FALSE,
    h = 0.1,
    n = 200
  ) +
  # Set alpha scaling with a minimum threshold
  scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
  # Use color scheme similar to leaflet's default heatmap
  scale_fill_gradientn(
    colors = c("#0000FF", "#00FFFF", "#00FF00", "#FFFF00", "#FF0000"),
    name = "Accident\nDensity"
  ) +
  coord_sf(
    xlim = c(-124.5, -114.5),
    ylim = c(32.5, 42.5)
  ) +
  theme_minimal() +
  labs(
    title = "Accident Density Heatmap in California (2016-2021)",
    x = NULL,
    y = NULL
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )
# Display the plot
print(accident_heatmap)

ggsave(filename = "imgs/accident_heatmap.png", 
       plot = accident_heatmap,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution

```

```{r Construction Heatmap static}
# Create heatmap of construction sites using ggplot2
construction_heatmap <- ggplot() +
  # Add California roads as base layer
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  # Create density heatmap
  stat_density_2d(
    data = df.const %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density)),
    geom = "tile", 
    contour = FALSE,
    alpha = 0.7,
    h = 0.1,  # Bandwidth
    n = 200   # Resolution
  ) +
  # Set alpha scaling with a minimum threshold
  scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
  # Use color scheme matching your leaflet construction map
  scale_fill_gradientn(
    colors = c("yellow", "orange", "red"),
    name = "Construction\nDensity"
  ) +
  # Set appropriate boundaries for California
  coord_sf(
    xlim = c(-124.5, -114.5),
    ylim = c(32.5, 42.5)
  ) +
  theme_minimal() +
  labs(
    title = "Construction Density Heatmap in California (2016-2021)",
    x = NULL,
    y = NULL
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )
# Display the plot
print(construction_heatmap)
```

## Heatmaps - interactive

```{r Interactive Maps I}
# For interactive maps (often better for large datasets)
# Accidents map
accident_map <- leaflet() %>%
  addTiles() %>%
  addHeatmap(data = df.acc.sf, 
             intensity = ~1,
             radius = 8, 
             blur = 10) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>% # Center to CA
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display maps
accident_map
```

```{r Interactive Maps II}
# Construction sites map with California focus and bounds constraints
construction_map <- leaflet() %>%
  addTiles() %>%
  addHeatmap(data = df.const.sf, 
             intensity = ~1,
             radius = 8, 
             blur = 10,
             gradient = c("yellow", "orange", "red")) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display the map
construction_map
```

We construct construction sites now also as line objects since they have a start and end point

```{r Construction linestrings}
# Convert construction data to linestring sf object
df.const.lines <- df.const %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng) & !is.na(End_Lat) & !is.na(End_Lng)) %>%
  mutate(geometry = pmap(list(Start_Lng, Start_Lat, End_Lng, End_Lat), 
                         function(start_lng, start_lat, end_lng, end_lat) {
                           coords <- matrix(c(start_lng, start_lat, 
                                             end_lng, end_lat), 
                                           ncol = 2, byrow = TRUE)
                           st_linestring(coords)
                         })) %>%
  st_sf(crs = 4326)
```

## Spatial operations in PostgreSQL

Since file sizes are so large and Spatial operations in R incredibly slow, we retreat to a PostgreSQL database for the following operations (PostGIS also uses the SF framework, thus R are almost identical to the following SQL functions used).

```{r DB ops}
# Create a connection to your PostgreSQL database
con <- dbConnect(
  Postgres(),
  dbname = "geospatial_final",
  host = "localhost",
  port = 5432,
  # Uncomment and add your credentials if needed
  user = "postgres",
  password = Sys.getenv("POSTGRES_PWD")
)

# Set connection parameters for performance optimization
dbExecute(con, "SET work_mem = '128MB'")           # Increase working memory
dbExecute(con, "SET maintenance_work_mem = '256MB'") # For index creation
dbExecute(con, "SET random_page_cost = 1.1")       # For SSD storage
dbExecute(con, "SET effective_cache_size = '2GB'") # Estimate of memory available
dbExecute(con, "SET max_parallel_workers_per_gather = 4") # Use multiple cores

# Check if PostGIS extension is enabled
postgis_check <- dbGetQuery(con, "SELECT PostGIS_version()")
cat("PostGIS version:", postgis_check[[1]], "\n")

# Check if SRID 3310 is available and add if not
dbExecute(con, "
  DO $$
  BEGIN
    IF NOT EXISTS (SELECT 1 FROM spatial_ref_sys WHERE srid = 3310) THEN
      INSERT INTO spatial_ref_sys (srid, auth_name, auth_srid, proj4text, srtext)
      VALUES (3310, 'EPSG', 3310, 
      '+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +datum=NAD83 +units=m +no_defs',
      'PROJCS[\"NAD83 / California Albers\",GEOGCS[\"NAD83\",DATUM[\"North_American_Datum_1983\",SPHEROID[\"GRS 1980\",6378137,298.257222101,AUTHORITY[\"EPSG\",\"7019\"]],AUTHORITY[\"EPSG\",\"6269\"]],PRIMEM[\"Greenwich\",0,AUTHORITY[\"EPSG\",\"8901\"]],UNIT[\"degree\",0.01745329251994328,AUTHORITY[\"EPSG\",\"9122\"]],AUTHORITY[\"EPSG\",\"4269\"]],PROJECTION[\"Albers_Conic_Equal_Area\"],PARAMETER[\"standard_parallel_1\",34],PARAMETER[\"standard_parallel_2\",40.5],PARAMETER[\"latitude_of_center\",0],PARAMETER[\"longitude_of_center\",-120],PARAMETER[\"false_easting\",0],PARAMETER[\"false_northing\",-4000000],UNIT[\"metre\",1,AUTHORITY[\"EPSG\",\"9001\"]],AUTHORITY[\"EPSG\",\"3310\"]]');
    END IF;
  END
  $$;
")

# Check if tables exist and only create them if they don't
# Begin transaction for data processing
dbExecute(con, "BEGIN")

# Check if accidents table exists, if not create it
has_accidents <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'accidents')")$exists
if (!has_accidents) {
  # Write accidents data to database
  sf::st_write(df.acc.sf, con, "accidents", 
               layer_options = c("GEOM_TYPE=GEOMETRY", 
                                 "GEOMETRY_NAME=geometry",
                                 "SRID=4326"))
  cat("Written accidents data to database\n")

  # Create spatial index for accidents
  dbExecute(con, "CREATE INDEX idx_accidents_geometry ON accidents USING GIST(geometry)")
  dbExecute(con, 'CREATE INDEX idx_accidents_start_time ON accidents("Start_Time");')
  dbExecute(con, "ANALYZE accidents")
  cat("Created spatial index for accidents table\n")
}

# Check if construction table exists, if not create it
has_construction <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction')")$exists

if (!has_construction) {
  # Write construction data to database
  sf::st_write(df.const.sf, con, "construction", 
               layer_options = c("GEOM_TYPE=GEOMETRY", 
                                 "GEOMETRY_NAME=geometry",
                                 "SRID=4326"))
  cat("Written construction data to database\n")

  # Create spatial index for construction
  dbExecute(con, "CREATE INDEX idx_construction_geometry ON construction USING GIST(geometry)")
  dbExecute(con, 'CREATE INDEX idx_construction_start_time ON construction("Start_Time");')
  dbExecute(con, 'CREATE INDEX idx_construction_end_time ON construction("End_Time");')
  dbExecute(con, "ANALYZE construction")
  cat("Created spatial index for construction table\n")
}

# Check if construction table with lines exists, if not create it
has_construction_lines <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_lines')")$exists

if (!has_construction_lines) {
  # Write construction data to database
  sf::st_write(df.const.lines, con, "construction_lines", 
               layer_options = c("GEOM_TYPE=GEOMETRY", 
                                 "GEOMETRY_NAME=geometry",
                                 "SRID=4326"))
  cat("Written construction_lines data to database\n")

  # Create spatial index for construction
  dbExecute(con, "CREATE INDEX idx_construction_lines_geometry ON construction_lines USING GIST(geometry)")
  dbExecute(con, 'CREATE INDEX idx_construction_lines_start_time ON construction_lines("Start_Time");')
  dbExecute(con, 'CREATE INDEX idx_construction_lines_end_time ON construction_lines("End_Time");')
  dbExecute(con, "ANALYZE construction_lines")
  cat("Created spatial index for construction_lines table\n")
}

# Commit transaction for base tables
dbExecute(con, "COMMIT")

#---------------------------------------------------------------------------
# STEP 1A: Create enhanced construction buffers table with time ranges
#---------------------------------------------------------------------------
has_construction_buffers <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_buffers')")$exists
if (!has_construction_buffers) {
  cat("Step 1A: Creating enhanced construction_buffers table with time ranges...\n")

  dbExecute(con, "BEGIN")

  # Create construction buffers table with appropriate projection and time ranges
  dbExecute(con, "
    CREATE TABLE construction_buffers AS
    SELECT 
      \"ID\" as construction_id,
      \"Start_Time\" as start_time,
      \"End_Time\" as end_time,
      ST_Transform(
        ST_Buffer(
          ST_Transform(geometry, 3310), -- Transform to California Albers for accurate distance
          300                           -- 300 meters buffer
        ), 
        4326                            -- Transform back to WGS 84 for joining
      ) AS geometry
    FROM construction
  ")

  # Add primary key and spatial index
  dbExecute(con, "ALTER TABLE construction_buffers ADD PRIMARY KEY (construction_id)")
  dbExecute(con, "CREATE INDEX idx_construction_buffers_geometry ON construction_buffers USING GIST(geometry)")
  dbExecute(con, "CREATE INDEX idx_construction_buffers_times ON construction_buffers(start_time, end_time)")
  dbExecute(con, "ANALYZE construction_buffers")

  dbExecute(con, "COMMIT")
  cat("Created enhanced construction_buffers table with spatial index and time ranges\n")
} else {
  cat("Construction buffers table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 1B: Create spatial matches table (using enhanced pre-computed buffers)
#---------------------------------------------------------------------------
has_spatial_matches <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'spatial_matches')")$exists
if (!has_spatial_matches) {
  cat("Step 1B: Creating spatial_matches table using enhanced pre-computed buffers...\n")

  dbExecute(con, "BEGIN")

  # Simplified join using pre-computed buffers with time information
  dbExecute(con, "
    CREATE TABLE spatial_matches AS
    SELECT
      a.\"ID\" AS accident_id,
      a.\"Severity\" AS accident_severity,
      a.\"Start_Time\" AS accident_time,
      b.construction_id
    FROM 
      accidents a
    JOIN 
      construction_buffers b ON 
        ST_Within(a.geometry, b.geometry) AND
        a.\"Start_Time\" BETWEEN b.start_time AND b.end_time
  ")

  # Create indices for efficient join operations in later steps
  dbExecute(con, "CREATE INDEX idx_spatial_accident ON spatial_matches(accident_id)")
  dbExecute(con, "CREATE INDEX idx_spatial_construction ON spatial_matches(construction_id)")
  dbExecute(con, "ANALYZE spatial_matches")

  dbExecute(con, "COMMIT")
  cat("Created spatial_matches table with combined temporal and spatial filtering\n")
} else {
  cat("Spatial matches table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 2: Count construction sites per accident (unchanged)
#---------------------------------------------------------------------------
has_construction_counts <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_counts')")$exists
if (!has_construction_counts) {
  cat("Step 2: Creating construction_counts table...\n")

  dbExecute(con, "BEGIN")

  dbExecute(con, "
    CREATE TABLE construction_counts AS
    SELECT
      accident_id,
      COUNT(*) AS nearby_construction_count
    FROM 
      spatial_matches
    GROUP BY
      accident_id
  ")

  # Create index for efficient joins in the final step
  dbExecute(con, "CREATE INDEX idx_construction_counts_accident ON construction_counts(accident_id)")
  dbExecute(con, "ANALYZE construction_counts")

  dbExecute(con, "COMMIT")
  cat("Created construction_counts table\n")
} else {
  cat("Construction counts table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 3: Create final analysis table (unchanged)
#---------------------------------------------------------------------------
has_accident_construction_analysis <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'accident_construction_analysis')")$exists
if (!has_accident_construction_analysis) {
  cat("Step 3: Creating final accident_construction_analysis table...\n")

  dbExecute(con, "BEGIN")

  dbExecute(con, "
    CREATE TABLE accident_construction_analysis AS
    WITH ranked_matches AS (
      SELECT
        sm.accident_id,
        sm.construction_id,
        sm.accident_severity,
        sm.accident_time,
        cc.nearby_construction_count,
        -- Rank construction sites for each accident (random selection if multiple)
        ROW_NUMBER() OVER (
          PARTITION BY sm.accident_id
          ORDER BY RANDOM()  -- Random selection ensures unbiased assignment
        ) AS rank
      FROM 
        spatial_matches sm
      JOIN
        construction_counts cc ON sm.accident_id = cc.accident_id
    )
    SELECT 
      accident_id,
      construction_id,
      'LINE' AS construction_type,  -- Indicates we're working with line geometries only
      accident_severity,            -- For severity analysis
      accident_time,                -- For temporal analysis
      1 AS temporal_overlap,        -- Always 1 due to our temporal filter in Step 1
      nearby_construction_count,    -- Number of construction sites affecting this accident
      1 AS is_near_active_construction  -- Always 1 due to our filtering steps
    FROM 
      ranked_matches
    WHERE 
      rank = 1  -- Select only one construction site per accident
  ")

  # Create indices for query optimization
  dbExecute(con, "CREATE INDEX idx_aca_accident_id ON accident_construction_analysis(accident_id)")
  dbExecute(con, "CREATE INDEX idx_aca_construction_id ON accident_construction_analysis(construction_id)")
  dbExecute(con, "CREATE INDEX idx_aca_count ON accident_construction_analysis(nearby_construction_count)")
  dbExecute(con, "ANALYZE accident_construction_analysis")

  dbExecute(con, "COMMIT")
  cat("Created accident_construction_analysis table\n")
} else {
  cat("Accident construction analysis table already exists, skipping creation\n")
}

# Read the data for construction zones with accident counts (for visualization)
accidents_per_zone <- st_read(con, query = "
  WITH construction_accident_counts AS (
    SELECT 
      aca.construction_id,
      COUNT(aca.accident_id) AS accident_count
    FROM 
      accident_construction_analysis aca
    GROUP BY 
      aca.construction_id
  )
  SELECT 
    cl.\"ID\" AS construction_id,
    cac.accident_count,
    cl.geometry
  FROM 
    construction_lines cl
  JOIN 
    construction_accident_counts cac ON cl.\"ID\" = cac.construction_id
  ORDER BY 
    cac.accident_count DESC
")

# Get statistics on construction zones
cat("Construction zones with accidents:", nrow(accidents_per_zone), "\n")
cat("Maximum accidents in a single construction zone:", max(accidents_per_zone$accident_count), "\n")
cat("Average accidents per construction zone:", round(mean(accidents_per_zone$accident_count), 2), "\n")

# Read accidents with their associated construction data
accidents_with_construction <- st_read(con, query = "
  SELECT 
    a.*,
    CASE WHEN aca.accident_id IS NOT NULL THEN 1 ELSE 0 END AS near_construction,
    aca.construction_id,
    aca.nearby_construction_count,
    aca.construction_type
  FROM 
    accidents a
  LEFT JOIN 
    accident_construction_analysis aca ON a.\"ID\" = aca.accident_id
")

# Get distribution of accident counts by construction density
construction_density_stats <- dbGetQuery(con, "
  SELECT 
    nearby_construction_count,
    COUNT(*) AS accident_count
  FROM 
    accident_construction_analysis
  GROUP BY 
    nearby_construction_count
  ORDER BY 
    nearby_construction_count
")

#---------------------------------------------------------------------------
# Report completion and verify table sizes
#---------------------------------------------------------------------------
cat("All tables are ready for analysis\n")

print(construction_density_stats)

# Total accidents summary - for comparison
total_accidents <- dbGetQuery(con, "SELECT COUNT(*) as count FROM accidents")$count
accidents_near_construction <- dbGetQuery(con, "SELECT COUNT(*) as count FROM accident_construction_analysis")$count

cat("Total accidents in dataset:", total_accidents, "\n")
cat("Accidents near active construction:", accidents_near_construction, "\n")
cat("Percentage of accidents near active construction: ", 
    round(accidents_near_construction/total_accidents*100, 2), "%\n")

#---------------------------------------------------------------------------
# STEP 1A-L: Create construction line buffers table with time ranges
#---------------------------------------------------------------------------
has_construction_line_buffers <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_line_buffers')")$exists
if (!has_construction_line_buffers) {
  cat("Step 1A-L: Creating construction_line_buffers table with time ranges...\n")

  dbExecute(con, "BEGIN")

  # Create line-based construction buffers with time information
  dbExecute(con, "
    CREATE TABLE construction_line_buffers AS
    SELECT 
      \"ID\" as construction_id,
      \"Start_Time\" as start_time,
      \"End_Time\" as end_time,
      ST_Transform(
        ST_Buffer(
          ST_Transform(geometry, 3310), -- Transform to California Albers for accurate distance
          300                           -- 300 meters buffer
        ), 
        4326                            -- Transform back to WGS 84 for joining
      ) AS geometry,
      ST_Length(ST_Transform(geometry, 3310)) AS segment_length_m -- Length in meters
    FROM construction_lines
  ")

  # Add primary key and spatial index
  dbExecute(con, "ALTER TABLE construction_line_buffers ADD PRIMARY KEY (construction_id)")
  dbExecute(con, "CREATE INDEX idx_construction_line_buffers_geometry ON construction_line_buffers USING GIST(geometry)")
  dbExecute(con, "CREATE INDEX idx_construction_line_buffers_times ON construction_line_buffers(start_time, end_time)")
  dbExecute(con, "ANALYZE construction_line_buffers")

  dbExecute(con, "COMMIT")
  cat("Created construction_line_buffers table with spatial index and time ranges\n")
} else {
  cat("Construction line buffers table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 1B-L: Create spatial matches table for line objects
#---------------------------------------------------------------------------
has_line_spatial_matches <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'line_spatial_matches')")$exists
if (!has_line_spatial_matches) {
  cat("Step 1B-L: Creating line_spatial_matches table using pre-computed line buffers...\n")

  dbExecute(con, "BEGIN")

  # Simplified join using pre-computed line buffers with time information
  dbExecute(con, "
    CREATE TABLE line_spatial_matches AS
    SELECT
      a.\"ID\" AS accident_id,
      a.\"Severity\" AS accident_severity,
      a.\"Start_Time\" AS accident_time,
      b.construction_id,
      b.segment_length_m
    FROM 
      accidents a
    JOIN 
      construction_line_buffers b ON 
        ST_Within(a.geometry, b.geometry) AND
        a.\"Start_Time\" BETWEEN b.start_time AND b.end_time
  ")

  # Create indices for efficient join operations
  dbExecute(con, "CREATE INDEX idx_line_spatial_accident ON line_spatial_matches(accident_id)")
  dbExecute(con, "CREATE INDEX idx_line_spatial_construction ON line_spatial_matches(construction_id)")
  dbExecute(con, "ANALYZE line_spatial_matches")

  dbExecute(con, "COMMIT")
  cat("Created line_spatial_matches table with combined temporal and spatial filtering\n")
} else {
  cat("Line spatial matches table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 2-L: Count construction line sites per accident
#---------------------------------------------------------------------------
has_line_construction_counts <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'line_construction_counts')")$exists
if (!has_line_construction_counts) {
  cat("Step 2-L: Creating line_construction_counts table...\n")

  dbExecute(con, "BEGIN")

  dbExecute(con, "
    CREATE TABLE line_construction_counts AS
    SELECT
      accident_id,
      COUNT(*) AS nearby_line_construction_count
    FROM 
      line_spatial_matches
    GROUP BY
      accident_id
  ")

  # Create index for efficient joins in the final step
  dbExecute(con, "CREATE INDEX idx_line_construction_counts_accident ON line_construction_counts(accident_id)")
  dbExecute(con, "ANALYZE line_construction_counts")

  dbExecute(con, "COMMIT")
  cat("Created line_construction_counts table\n")
} else {
  cat("Line construction counts table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 3-L: Create final line-based analysis table
#---------------------------------------------------------------------------
has_line_accident_construction_analysis <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'line_accident_construction_analysis')")$exists
if (!has_line_accident_construction_analysis) {
  cat("Step 3-L: Creating final line_accident_construction_analysis table...\n")

  dbExecute(con, "BEGIN")

  dbExecute(con, "
    CREATE TABLE line_accident_construction_analysis AS
    WITH ranked_matches AS (
      SELECT
        sm.accident_id,
        sm.construction_id,
        sm.accident_severity,
        sm.accident_time,
        sm.segment_length_m,
        cc.nearby_line_construction_count,
        -- Rank construction sites for each accident (random selection if multiple)
        ROW_NUMBER() OVER (
          PARTITION BY sm.accident_id
          ORDER BY RANDOM()  -- Random selection ensures unbiased assignment
        ) AS rank
      FROM 
        line_spatial_matches sm
      JOIN
        line_construction_counts cc ON sm.accident_id = cc.accident_id
    )
    SELECT 
      accident_id,
      construction_id,
      accident_severity,
      accident_time,
      segment_length_m,
      nearby_line_construction_count,
      1 AS temporal_overlap,           -- Always 1 due to our temporal filter in Step 1
      1 AS is_near_active_construction -- Always 1 due to our filtering steps
    FROM 
      ranked_matches
    WHERE 
      rank = 1  -- Select only one construction site per accident
  ")

  # Create indices for query optimization
  dbExecute(con, "CREATE INDEX idx_laca_accident_id ON line_accident_construction_analysis(accident_id)")
  dbExecute(con, "CREATE INDEX idx_laca_construction_id ON line_accident_construction_analysis(construction_id)")
  dbExecute(con, "CREATE INDEX idx_laca_count ON line_accident_construction_analysis(nearby_line_construction_count)")
  dbExecute(con, "ANALYZE line_accident_construction_analysis")

  dbExecute(con, "COMMIT")
  cat("Created line_accident_construction_analysis table\n")
} else {
  cat("Line accident construction analysis table already exists, skipping creation\n")
}

#---------------------------------------------------------------------------
# STEP 4-L: Calculate accident rates by segment length
#---------------------------------------------------------------------------
has_accident_rates <- dbGetQuery(con, "SELECT EXISTS (SELECT 1 FROM information_schema.tables WHERE table_schema = 'public' AND table_name = 'construction_segment_accident_rates')")$exists
if (!has_accident_rates) {
  cat("Step 4-L: Creating construction_segment_accident_rates table...\n")

  dbExecute(con, "BEGIN")

  dbExecute(con, "
    CREATE TABLE construction_segment_accident_rates AS
    SELECT 
      cl.\"ID\" AS construction_id,
      cl.geometry,
      COUNT(laca.accident_id) AS accident_count,
      MAX(laca.segment_length_m) AS segment_length_m,
      COUNT(laca.accident_id) / (MAX(laca.segment_length_m) / 1000.0) AS accidents_per_km
    FROM 
      construction_lines cl
    LEFT JOIN 
      line_accident_construction_analysis laca ON cl.\"ID\" = laca.construction_id
    GROUP BY 
      cl.\"ID\", cl.geometry
  ")

  # Create spatial index
  dbExecute(con, "CREATE INDEX idx_accident_rates_geometry ON construction_segment_accident_rates USING GIST(geometry)")
  dbExecute(con, "ANALYZE construction_segment_accident_rates")

  dbExecute(con, "COMMIT")
  cat("Created construction_segment_accident_rates table\n")
} else {
  cat("Construction segment accident rates table already exists, skipping creation\n")
}

# Read accident rates by construction segment
accident_rates_by_segment <- st_read(con, query = "
  SELECT 
    construction_id,
    accident_count,
    segment_length_m,
    accidents_per_km,
    geometry
  FROM 
    construction_segment_accident_rates
  WHERE 
    segment_length_m > 0  -- Avoid division by zero
  ORDER BY 
    accidents_per_km DESC
")

# Display summary statistics
cat("Construction segments analyzed:", nrow(accident_rates_by_segment), "\n")
cat("Maximum accidents per km:", max(accident_rates_by_segment$accidents_per_km, na.rm = TRUE), "\n")
cat("Average accidents per km:", mean(accident_rates_by_segment$accidents_per_km, na.rm = TRUE), "\n")
cat("Median segment length (m):", median(accident_rates_by_segment$segment_length_m, na.rm = TRUE), "\n")

# Severity analysis near line construction
severity_by_construction <- dbGetQuery(con, "
  SELECT 
    laca.accident_severity, 
    COUNT(*) AS count
  FROM 
    line_accident_construction_analysis laca
  GROUP BY 
    laca.accident_severity
  ORDER BY 
    laca.accident_severity
")

# Comparison of accident severity near vs. away from construction
severity_comparison <- dbGetQuery(con, "
  WITH construction_accidents AS (
    SELECT 
      a.\"ID\", 
      a.\"Severity\", 
      'Near construction' AS location
    FROM 
      accidents a
    JOIN 
      line_accident_construction_analysis laca ON a.\"ID\" = laca.accident_id
  ),
  non_construction_accidents AS (
    SELECT 
      a.\"ID\", 
      a.\"Severity\", 
      'Away from construction' AS location
    FROM 
      accidents a
    LEFT JOIN 
      line_accident_construction_analysis laca ON a.\"ID\" = laca.accident_id
    WHERE 
      laca.accident_id IS NULL
  )
  SELECT 
    \"Severity\", 
    location, 
    COUNT(*) AS count
  FROM (
    SELECT * FROM construction_accidents
    UNION ALL
    SELECT * FROM non_construction_accidents
  ) combined
  GROUP BY 
    \"Severity\", location
  ORDER BY 
    \"Severity\", location
")

# Time-based analysis
temporal_analysis <- dbGetQuery(con, "
  SELECT 
    DATE_TRUNC('month', laca.accident_time) AS month,
    COUNT(*) AS accident_count
  FROM 
    line_accident_construction_analysis laca
  GROUP BY 
    DATE_TRUNC('month', laca.accident_time)
  ORDER BY 
    month
")

# Disconnect from database
dbDisconnect(con)
```

## Buffer zone analysis

In theory, the commented code below would achieve the same as our operation PostGIS but way slower. We will not run it here, but keep it to demonstrate that we could implement in R as well.

```{r Buffer Zone Analysis}
# Set up parallel processing
#future::plan(future::multisession, workers = parallel::detectCores() - 1)

# Set sf to use parallel processing
#sf_use_s2(FALSE)  # Disable S2 spherical geometry
#cores <- parallel::detectCores() - 1

# Process each dataset with batching
#cat("Creating point buffers...\n")
#const_buffers <- create_buffers_in_batches(df.const.sf, buffer_dist = 1000, batch_size = 500)

# Compute accidents within construction zones
#accidents_in_construction <- df.acc.sf %>%
#  st_join(const_buffers, join = st_within, prepared = TRUE)

# Return to sequential processing
#future::plan(future::sequential)
#sf_use_s2(TRUE)  # Re-enable S2 spherical geometry

# Count accidents per construction zone
#accidents_per_zone <- accidents_with_construction %>%
#  group_by(ID) %>%  # Grouping by the ID column
#  summarize(accident_count = n())

accidents_per_zone <- accidents_per_zone %>% drop_na()

# Visualize construction zones with accident counts
# First, store our plot in a variable
accident_zones_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  # Arrange data in descending order so high frequency zones are plotted on top
  geom_sf(data = accidents_per_zone %>% arrange(accident_count), 
          aes(color = accident_count), size = 1.5) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Accident Counts within Construction Zones (2021)")

print(accident_zones_plot)

# Then save it using ggsave
ggsave(filename = "imgs/accident_counts_construction_zones.png", 
       plot = accident_zones_plot,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution
```

## Advanced construction sit analysis (line objects):

```{r Maps with Construction Lines - static}
# Static map with construction line segments
ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.const.lines %>% filter(State == "CA"), 
          color = "orange", size = 1) +
  theme_minimal() +
  labs(title = "US Construction Segments (2021)")
```

```{r Maps with Construction Lines - dynamic}
# Interactive map with construction segments focused on California
construction_segment_map <- leaflet() %>%
  addTiles() %>%
  addPolylines(data = df.const.lines, 
               color = "orange",
               weight = 3,
               opacity = 0.7,
               popup = ~paste("ID:", ID, "<br>Duration:", duration)) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display the map
construction_segment_map
```

Again, we execute computationally heavy operations in PostGIS but have the commented R code below to showcase a theoretical, but slow implementatiojn in R.

```{r Construction Impact Analysis with Buffers}
# Set up parallel processing
#future::plan(future::multisession, workers = parallel::detectCores() - 1)

# Set sf to use parallel processing
#f_use_s2(FALSE)  # Disable S2 spherical geometry
#cores <- parallel::detectCores() - 1

#cat("Creating line segment buffers...\n")
#const_segment_buffers <- create_buffers_in_batches(df.const.lines, buffer_dist = 1000, batch_size = 250)

# Find accidents near construction segments
#accidents_near_construction <- df.acc.sf %>%
#  st_join(const_segment_buffers, join = st_intersects, prepared = TRUE)

# Return to sequential processing
#future::plan(future::sequential)
#sf_use_s2(TRUE)  # Re-enable S2 spherical geometry

# Analyze accident rate per construction segment length
#accidents_per_segment <- accidents_near_construction %>%
#  group_by(ID) %>%  # Assuming there's an ID column for construction segments
#  summarize(
#    accident_count = n(),
#    segment_length = as.numeric(st_length(geometry[1])),  # Length in meters
#    accidents_per_km = accident_count / (segment_length/1000)
#  )

# Visualize segments by accident rate
ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = accident_rates_by_segment, aes(color = accidents_per_km), size = 1.5) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Accident Rate per Construction Segment (2016-2021)",
       color = "Accidents/km")
```

```{r}
# Reestablish connection to PostgreSQL database
con <- dbConnect(
  Postgres(),
  dbname = "geospatial_final",
  host = "localhost",
  port = 5432,
  user = "postgres",
  password = Sys.getenv("POSTGRES_PWD")
)

# Temporal analysis - construction impact over time
accidents_by_time <- dbGetQuery(con, "
  SELECT 
    DATE_TRUNC('month', accident_time) AS month,
    COUNT(*) AS accident_count
  FROM 
    line_accident_construction_analysis
  GROUP BY 
    DATE_TRUNC('month', accident_time)
  ORDER BY 
    month
") %>%
  mutate(month = as.Date(month))  # Convert PostgreSQL timestamp to R Date

# Get severity comparison data
severity_comparison <- dbGetQuery(con, "
  WITH construction_accidents AS (
    SELECT 
      a.\"Severity\", 
      'Near construction' AS location
    FROM 
      accidents a
    JOIN 
      line_accident_construction_analysis laca ON a.\"ID\" = laca.accident_id
  ),
  non_construction_accidents AS (
    SELECT 
      a.\"Severity\", 
      'Away from construction' AS location
    FROM 
      accidents a
    LEFT JOIN 
      line_accident_construction_analysis laca ON a.\"ID\" = laca.accident_id
    WHERE 
      laca.accident_id IS NULL
  )
  SELECT 
    \"Severity\", 
    location, 
    COUNT(*) AS count
  FROM (
    SELECT * FROM construction_accidents
    UNION ALL
    SELECT * FROM non_construction_accidents
  ) combined
  GROUP BY 
    \"Severity\", location
  ORDER BY 
    \"Severity\", location
")

# Disconnect from database
dbDisconnect(con)
```

```{r Temporal Analysis and Severity Comparison with Buffers}
# Visualize temporal analysis
ggplot(accidents_by_time, aes(x = month, y = accident_count)) +
  geom_line() +
  geom_point() +
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Accidents Near Construction Over Time",
       x = "Month",
       y = "Accident Count")

# Visualize severity comparison
ggplot(severity_comparison, aes(x = Severity, y = count, fill = location)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(title = "Accident Severity: Near vs. Away from Construction",
       x = "Severity Level",
       y = "Count",
       fill = "Location")
```