### 2025-03-26 final/final_MoritzPeistNoemiLucchiSimonVellin.Rmd

---
title: "California Road Construction Safety Analysis: A Spatial Econometric Approach"
author:
- Moritz Peist
- Noemi Lucchi
- Simon Vellin
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
    code_folding: show
    fig_caption: true
    number_sections: true
    progress: false
    warning: false
---

```{css, echo=FALSE}
h1.title {
  text-align: center;
}

h4.author, h4.date {
  text-align: center;
}

.author {
  text-align: center;
}

.date {
  text-align: center;
}

h3.subtitle {
  text-align: center;
  font-style: italic;
  margin-top: -10px;
  color: #606060;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE, 
  message = FALSE,
  fig.width = 8,
  fig.height = 5,
  out.width = "100%",
  fig.align = "center",
  comment = "#>"
)

# Load required libraries
library(data.table)      # For faster data manipulation
library(tidyverse)       # For data manipulation and visualization
library(sf)              # For spatial data handling
library(leaflet)         # For interactive maps
library(leaflet.extras)  # For additional leaflet features
library(mapview)         # For easier map visualization
library(tmap)            # For thematic maps
library(tigris)          # For US road networks
library(fixest)          # For fixed effects models
library(spdep)           # For spatial weights and tests
library(splm)            # For spatial panel models
library(RColorBrewer)    # For color palettes
library(patchwork)       # For combining plots
library(kableExtra)      # For nice tables
library(lubridate)       # For date manipulation
library(viridis)         # For color palettes

# Create directories if they don't exist
if (!dir.exists("./data/tigris")) {
  dir.create("./data/tigris", recursive = TRUE)
}

# Set custom cache directory
options(tigris_cache_dir = "./data/tigris")
# Configure tigris to use caching
options(tigris_use_cache = TRUE)
```

<div style="text-align: center;">
  <img src="../Barcelona_School_of_Economics_logo.svg" alt="BSE logo" style="width: 200px;">
</div>

# Abstract

This study examines the causal relationship between road construction activities and traffic accident rates in California using high-resolution spatial data from 2021. Implementing a difference-in-differences framework with spatial controls, we analyze how construction zones affect accident frequency and severity. Our findings indicate a significant relationship between active construction and accident occurrence, with heterogeneous effects across urban and rural environments. Spatial spillover analysis reveals that accident impacts extend beyond immediate construction zones, with pronounced effects in areas surrounding construction sites. This research provides evidence-based insights for transportation safety planning and targeted mitigation strategies.

# Introduction

Road construction is a necessary yet potentially disruptive component of transportation infrastructure maintenance and development. While essential for long-term safety improvements, construction activities create temporary changes in traffic patterns, road conditions, and driver behavior that may increase accident risk. Understanding the causal relationship between construction activities and traffic safety has important implications for construction scheduling, resource allocation, traffic management strategies, and policy development.

Despite the importance of this relationship, rigorous causal evidence on how road construction affects traffic accidents remains limited, particularly with respect to potential heterogeneity between urban and rural environments. This study addresses this gap by leveraging comprehensive spatial data on road construction activities and traffic accidents in California during 2021 to answer the following research questions:

1. How does road construction causally impact traffic accident rates?
2. Do these effects systematically differ between urban and rural environments?
3. Are there spatial spillover effects beyond the immediate construction zones?

The findings of this research have direct implications for transportation safety planning, particularly regarding how resources should be allocated for traffic management during construction activities across different environments.

# Literature Review

To establish a contextual foundation for our study, we reviewed existing literature on the relationship between road construction and traffic accidents, focusing on studies that use spatial or causal inference approaches.

## 2.1 ["Effects of truck traffic on crash injury severity on rural highways in Wyoming using Bayesian binary logit models" by Ahmed et al. (2018)](https://www.sciencedirect.com/science/article/pii/S0001457518301507)

This paper explores the impact of truck traffic and roadway geometry on crash frequency within highway work zones, using a case study from Wyoming spanning 2012–2016.

Using crash data and statistical modeling, the authors assess how construction-related factors, like lane closures and heavy vehicle presence, elevate accident risks. Their paper reveals that increased truck volumes and sharper geometric features significantly amplify crash rates, with work zones acting as hotspots for such incidents. While the paper employs logistic regression rather than causal inference techniques, it underscores the real-world relevance of construction disruptions on traffic safety.

This work provides a practical foundation for our study, highlighting the need to isolate causal effects of construction using advanced methods. By focusing on observable crash patterns, Ahmed et al. set the stage for deeper methodological exploration in our difference-in-differences framework.

## 2.2 ["Comparison of characteristics between fatal and injury accidents in the highway construction zones" by Li and Bai (2008)]( https://www.sciencedirect.com/science/article/pii/S0925753507001014)

This paper investigates the characteristics of accidents in highway construction zones from a statistical perspective, comparing fatal and injury crashes using data from Utah between 1992 and 2004.

Using traffic volume, construction duration, and time of day, their results show that construction zones markedly increase accident likelihood, with fatal crashes tied to higher speeds and nighttime conditions. Although the study relies on descriptive statistics and regression rather than causal frameworks, it offers critical insights into how construction alters road safety dynamics.

## Accident Risk in Highway Construction Zones

The relationship between construction activities and traffic accidents has been investigated in several studies. Ahmed et al. (2018) explored the impact of truck traffic and roadway geometry on crash frequency within highway work zones in Wyoming. Their analysis revealed that increased truck volumes and sharper geometric features significantly amplified crash rates, with work zones acting as hotspots for accidents. While their study employed logistic regression rather than causal inference techniques, it highlighted the real-world relevance of construction disruptions on traffic safety.

Li and Bai (2008) compared the characteristics of fatal and injury crashes in highway construction zones using data from Utah (1992-2004). Their results showed that construction zones markedly increase accident likelihood, with fatal crashes tied to higher speeds and nighttime conditions. Although they relied on descriptive statistics rather than causal frameworks, their work established a clear link between construction activities and accident rates.

The literature generally suggests that construction zones are associated with increased accident risk, but most studies have not employed causal inference methods to isolate this effect from confounding factors. Additionally, few studies have examined the spatial heterogeneity of construction effects on accidents, particularly regarding urban-rural differences. Our study addresses these gaps by employing a difference-in-differences framework with spatial controls to identify the causal impact of construction activities on accident rates.

# Data and Methods

## Data Sources

This study integrates multiple large-scale datasets to create a comprehensive spatial-temporal framework for analysis:

**Primary Datasets:**

  1. [**US Road Construction and Closures Dataset (2021)**](https://www.kaggle.com/datasets/sobhanmoosavi/us-road-construction-and-closures) - Contains detailed records of construction activities in California with coordinates, start/end times, and affected distances
  2. [**US Traffic Accidents Dataset (2021)**](https://www.kaggle.com/datasets/sobhanmoosavi/us-accidents) - Comprehensive accident records with precise locations, timestamps, and severity measures
  3. [**Caltrans State Highway Network**](https://catalog.data.gov/dataset/traffic-volumes-aadt-ee8d6/resource/904a8300-9b57-4765-8ed7-cca0c83cdede) - Annual Average Daily Traffic (AADT) data for normalizing accident rates by traffic volume
  4. **US Census Bureau Urban Areas** - Official urban/rural classifications for spatial heterogeneity analysis

For both the accident and construction datasets, the data was collected using multiple APIs such as the US and state departments of transportation, law enforcement agencies, traffic cameras, and traffic sensors within the road networks.

While our analysis uses 2021 accident and construction data, we integrated Annual Average Daily Traffic (AADT) data from Caltrans to normalize accident rates. The AADT dataset provides a comprehensive traffic volume baseline that reflects typical traffic patterns, mitigating potential short-term fluctuations. Though the AADT data may include projections or estimates from surrounding years, it represents the most reliable and granular traffic volume information available, enabling more precise exposure-adjusted accident rate calculations.

## Data Preparation and Exploration

We focused our analysis on California road construction and accidents from 2021, creating a spatial-temporal framework that allows us to examine the causal relationship between construction activities and accidents.

```{r data_loading}
# Load Caltrans State Highway Network
aadt <- st_read("data/Traffic_Volumes_AADT/Traffic_Volumes_AADT.shp")

# Load and filter accidents data
df.acc <- fread("data/us_accidents/US_accidents_March23.csv")[
  # Filter date range of 2021
  lubridate::year(as.Date(Start_Time)) == 2021 & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = data.table::year(Start_Time),
  quarter = data.table::quarter(Start_Time),
  month = data.table::month(Start_Time),
  # Calculate duration
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)] %>% 
  as_tibble()

# Load and filter construction data
df.const <- fread("data/us_constructions/US_constructions_Dec21.csv")[
  # Filter date range of 2021
  lubridate::year(as.Date(Start_Time)) == 2021 & 
  # And California
  State == "CA"
][, `:=`(
  # Add year, quarter, month columns
  year = year(Start_Time),
  quarter = quarter(Start_Time),
  month = month(Start_Time),
  # Calculate duration
  duration = as.numeric(difftime(End_Time, Start_Time, units = "days"))
)][
  duration > 1  # Filter out constructions lasting less than a day
] %>% 
  as_tibble()

# Load California road network for mapping
ca_roads <- tigris::primary_roads("CA", year = 2021, refresh=TRUE) %>%
  st_transform(4326)
```

```{r data_summary}
# Calculate and display actual data summary
accidents_count <- nrow(df.acc)
construction_count <- nrow(df.const)
aadt_count <- nrow(aadt)

cat("Summary of datasets:\n")
cat("- Accidents dataset:", accidents_count, "records\n")
cat("- Construction dataset:", construction_count, "records\n")
cat("- AADT dataset:", aadt_count, "records\n")
```

### Temporal Patterns

```{r temporal_patterns}
# Temporal patterns visualization using actual data
acc_monthly <- df.acc %>%
  mutate(month = floor_date(as.Date(Start_Time), "month")) %>%
  count(month)

const_monthly <- df.const %>%
  mutate(month = floor_date(as.Date(Start_Time), "month")) %>%
  count(month)

# Plot monthly patterns
p1 <- ggplot(acc_monthly, aes(x = month, y = n)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "blue", size = 3) +
  labs(title = "Monthly Accidents in California (2021)",
       x = "Month", y = "Count") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

p2 <- ggplot(const_monthly, aes(x = month, y = n)) +
  geom_line(color = "red", size = 1) +
  geom_point(color = "red", size = 3) +
  labs(title = "New Construction Projects in California (2021)",
       x = "Month", y = "Count") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

p1 / p2
```

The temporal distribution reveals distinct patterns for accidents and construction projects. Accident rates peak in winter months (January and December) and fall (October), while construction activity is highest in spring (March-May). This inverse relationship suggests seasonal factors may influence both construction planning and accident occurrence.

### Spatial Distribution

Let's examine the spatial distribution of accidents and construction sites across California.

```{r spatial_processing}
# Create spatial objects from actual data
if (!inherits(df.acc, "sf")) {
  df.acc.sf <- df.acc %>%
    filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
    st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326) %>%
    st_transform(crs = 3310)  # CA Albers for accurate distance
} else {
  df.acc.sf <- df.acc
}

if (!inherits(df.const, "sf")) {
  df.const.sf <- df.const %>%
    filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
    st_as_sf(coords = c("Start_Lng", "Start_Lat"), crs = 4326) %>%
    st_transform(crs = 3310)
} else {
  df.const.sf <- df.const
}
```

```{r static_maps}
# Create a subsample for visualization (plotting all points would be too dense)
set.seed(123)
acc_sample <- df.acc.sf %>%
  sample_n(min(10000, nrow(.)))

const_sample <- df.const.sf %>%
  sample_n(min(5000, nrow(.)))

# Static map for Accidents with month coloring
ca_acc_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = acc_sample, 
          aes(color = factor(month)), alpha = 0.5, size = 0.5) +
  scale_color_viridis_d(name = "Month", option = "plasma") +
  coord_sf(
    xlim = c(-124.5, -114.5),
    ylim = c(32.5, 42.5)
  ) +
  theme_minimal() +
  labs(title = "California Road Accidents (2021)") +
  theme(legend.position = "bottom")

# Construction sites map with severity coloring
ca_const_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = const_sample, 
          aes(color = factor(Severity)), alpha = 0.7, size = 0.5) +
  scale_color_viridis_d(name = "Severity", option = "plasma") +
  coord_sf(
    xlim = c(-124.5, -114.5),
    ylim = c(32.5, 42.5)
  ) +
  theme_minimal() +
  labs(title = "California Construction Sites by Severity (2021)") +
  theme(legend.position = "bottom")

# Display the plots
ca_acc_plot
ca_const_plot
```

### Density Heatmaps

To better visualize the spatial concentration, we created heatmaps for both accidents and construction sites.

```{r heatmaps}
# Sample data for creating heatmaps (using smaller sample for performance)
set.seed(123)
acc_sample_for_heatmap <- df.acc %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  sample_n(min(50000, nrow(.)))

const_sample_for_heatmap <- df.const %>%
  filter(!is.na(Start_Lat) & !is.na(Start_Lng)) %>%
  sample_n(min(10000, nrow(.)))

# Create heatmap of accidents using ggplot2
accident_heatmap <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  stat_density_2d(
    data = acc_sample_for_heatmap,
    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density), alpha = after_stat(density)),
    geom = "tile", 
    contour = FALSE,
    h = 0.2,  # Increased bandwidth for better visualization
    n = 150
  ) +
  scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
  scale_fill_gradientn(
    colors = c("#0000FF", "#00FFFF", "#00FF00", "#FFFF00", "#FF0000"),
    name = "Accident\nDensity"
  ) +
  coord_sf(
    xlim = c(-124.5, -114.5),
    ylim = c(32.5, 42.5)
  ) +
  theme_minimal() +
  labs(
    title = "Accident Density Heatmap in California (2021)",
    x = NULL,
    y = NULL
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )

# Construction sites heatmap
construction_heatmap <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  stat_density_2d(
    data = const_sample_for_heatmap,
    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density)),
    geom = "tile", 
    contour = FALSE,
    alpha = 0.7,
    h = 0.2,
    n = 150
  ) +
  scale_fill_gradientn(
    colors = c("yellow", "orange", "red"),
    name = "Construction\nDensity",
    na.value = NA,  # Makes low-density areas transparent
    limits = c(0.001, NA)  # Sets minimum value to show (adjust as needed)
  ) +
  coord_sf(
    xlim = c(-124.5, -114.5),
    ylim = c(32.5, 42.5)
  ) +
  theme_minimal() +
  labs(
    title = "Construction Density Heatmap in California (2021)",
    x = NULL,
    y = NULL
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )

# Display the heatmaps
accident_heatmap
construction_heatmap
```

The heatmaps confirm the spatial concentration of both accidents and construction activities in California's major urban areas. The similar patterns suggest a potential relationship, but we need to control for traffic volume differences and establish causality.

### Urban-Rural Classification

We classified construction sites as urban or rural based on their intersection with US Census Bureau urban areas. This classification is crucial for our heterogeneity analysis.

```{r urban_rural}
# Download urban areas for California
ca_urban_areas <- tigris::urban_areas(cb = TRUE, year = 2019) %>%
  st_transform(st_crs(df.const.sf)) %>%
  filter(grepl(", CA$", NAME10))

# Classify construction sites as urban or rural
const_urban <- st_join(
  df.const.sf,
  ca_urban_areas %>% select(urban_name = NAME10),
  join = st_intersects
)

const_urban <- const_urban %>%
  mutate(is_urban = !is.na(urban_name))
```

```{r urban_rural_table}
# Calculate actual urban-rural distribution
urban_rural_df <- const_urban %>%
  st_drop_geometry() %>%
  count(is_urban) %>%
  mutate(percentage = n / sum(n) * 100)

kableExtra::kable(urban_rural_df, 
                 col.names = c("Urban Area", "Count", "Percentage (%)"),
                 caption = "Distribution of Construction Sites by Urban/Rural Classification",
                 booktabs = TRUE,
                 digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),
                           full_width = FALSE) %>%
  kableExtra::row_spec(0, bold = TRUE)
```

Our analysis of the urban-rural breakdown is critical for understanding the heterogeneous effects of construction on accident rates.

### Integration with Traffic Volume Data

To control for exposure to traffic, we integrated Annual Average Daily Traffic (AADT) data from Caltrans, allowing us to normalize accident counts by traffic volume.

```{r traffic_volume}
# Transform AADT data to ensure matching CRS
aadt_sf <- st_transform(aadt, st_crs(df.const.sf))

# Convert AADT columns to numeric
aadt_sf <- aadt_sf %>%
  mutate(
    BACK_AADT_num = as.numeric(as.character(BACK_AADT)),
    AHEAD_AADT_num = as.numeric(as.character(AHEAD_AADT))
  )

# Find the nearest AADT point for each construction buffer
nearest_indices <- st_nearest_feature(df.const.sf, aadt_sf)
nearest_distances <- st_distance(df.const.sf, aadt_sf[nearest_indices,], by_element = TRUE)

# Create a dataframe with buffer IDs and nearest AADT values
buffer_aadt <- df.const.sf %>%
  mutate(buffer_id = paste0("c", row_number())) %>%
  select(buffer_id) %>%
  bind_cols(
    aadt_sf[nearest_indices, ] %>%
    select(BACK_AADT_num, AHEAD_AADT_num)
  ) %>%
  mutate(
    distance_to_nearest = as.numeric(nearest_distances),
    traffic_volume = rowMeans(cbind(BACK_AADT_num, AHEAD_AADT_num), na.rm = TRUE)
  ) %>%
  st_drop_geometry() %>%
  select(buffer_id, traffic_volume, distance_to_nearest)

# Summary of traffic volumes
summary(buffer_aadt$traffic_volume)
```

```{r traffic_distribution}
# Plot actual traffic volume distribution
ggplot(buffer_aadt, aes(x = traffic_volume)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "black", alpha = 0.8) +
  labs(
    title = "Distribution of Annual Average Daily Traffic (AADT)",
    x = "Traffic Volume (vehicles per day)",
    y = "Count"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

# Log-scale for better visualization of the distribution
ggplot(buffer_aadt, aes(x = traffic_volume)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "black", alpha = 0.8) +
  scale_x_log10() +
  labs(
    title = "Distribution of Annual Average Daily Traffic (AADT) - Log Scale",
    x = "Traffic Volume (vehicles per day) - Log Scale",
    y = "Count"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

The traffic volume distribution shows a pattern likely reflecting the differences between urban and rural areas. Urban areas typically have higher traffic volumes, which could influence both construction planning and accident rates.

## Construction Buffer Analysis

To establish the spatial relationship between accidents and construction activities, we created buffer zones around construction locations and analyzed accidents within these zones.

```{r construction_buffers}
# Create construction buffers (500m)
const_buffers <- const_urban %>%
  mutate(buffer_id = paste0("c", row_number())) %>%
  st_buffer(dist = 500)  # 500m buffer

# Add buffer metadata
const_buffers <- const_buffers %>%
  mutate(
    start_date = as.Date(Start_Time),
    end_date = as.Date(End_Time),
    construction_duration = as.numeric(difftime(end_date, start_date, units = "days"))
  )

# Join accidents to construction buffers - this operation can be computationally intensive
# For very large datasets, consider using a spatial database or partitioning the data
accidents_in_buffers <- st_join(
  df.acc.sf %>% 
    select(accident_id = ID, accident_date = Start_Time, Severity),
  const_buffers %>% 
    select(buffer_id, construction_id = ID, start_date, end_date, is_urban),
  join = st_intersects
)

# Classify accidents by timing relative to construction
accidents_in_buffers <- accidents_in_buffers %>%
  mutate(
    accident_date = as.Date(accident_date),
    time_period = case_when(
      is.na(start_date) ~ NA_character_,
      accident_date < start_date ~ "before",
      accident_date > end_date ~ "after",
      TRUE ~ "during"
    ),
    month_year = floor_date(accident_date, "month")
  )

# Display the first few rows of the joined data
head(accidents_in_buffers %>% st_drop_geometry())
```

```{r period_counts_table}
# Count accidents by construction period from actual data
period_df <- accidents_in_buffers %>%
  st_drop_geometry() %>%
  filter(!is.na(time_period)) %>%
  count(time_period) %>%
  mutate(percentage = n / sum(n) * 100)

kableExtra::kable(period_df, 
                 col.names = c("Construction Period", "Count", "Percentage (%)"),
                 caption = "Distribution of Accidents by Construction Period",
                 booktabs = TRUE,
                 digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),
                           full_width = FALSE) %>%
  kableExtra::row_spec(0, bold = TRUE)
```

When examining accidents by their timing relative to construction, this distribution provides initial evidence that construction activities may impact accident patterns, but we need more rigorous analysis to establish causality.

```{r buffer_visualization}
# Extract LA area data for visualization
la_bbox <- st_bbox(c(xmin = -118.75, ymin = 33.65, xmax = -117.85, ymax = 34.35), crs = st_crs(4326))
la_bbox_sf <- st_as_sfc(la_bbox)

# Filter construction buffers and accidents for LA area
la_buffers <- const_buffers %>%
  st_transform(4326) %>%
  st_filter(la_bbox_sf)

# Sample if there are too many for visualization
if(nrow(la_buffers) > 50) {
  set.seed(123)
  la_buffers <- la_buffers %>% sample_n(50)
}

la_accidents <- df.acc.sf %>%
  st_transform(4326) %>%
  st_filter(la_bbox_sf)

# Sample if there are too many for visualization
if(nrow(la_accidents) > 500) {
  set.seed(123)
  la_accidents <- la_accidents %>% sample_n(500)
}

# Create LA outline
la_outline_sf <- st_as_sf(st_as_sfc(la_bbox))

# Plot
ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = la_buffers, fill = "red", alpha = 0.3, color = "red") +
  geom_sf(data = st_centroid(la_buffers), color = "darkred", size = 2) +
  geom_sf(data = la_accidents, color = "blue", size = 1, alpha = 0.6) +
  labs(
    title = "Construction Buffers and Accidents in Los Angeles Area",
    subtitle = "Red areas: 500m construction buffers, Blue points: accidents"
  ) +
  coord_sf(
    xlim = c(-118.75, -117.85), 
    ylim = c(33.65, 34.35)
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

The visualization illustrates the spatial relationship between construction buffers (red areas) and accident locations (blue points). Many accidents occur within or near construction zones, suggesting a potential spatial correlation.

## Methodology

Our analysis employs a difference-in-differences (DiD) framework augmented with spatial controls to identify the causal impact of construction activities on traffic accidents. This approach allows us to address potential endogeneity concerns by controlling for time-invariant location characteristics and time-specific effects.

### Panel Dataset Construction

We created a panel dataset that tracks accidents in each buffer zone over time, with indicators for pre-construction, during-construction, and post-construction periods.

```{r panel_creation}
# Generate all buffer-month combinations
all_months <- seq(as.Date("2021-01-01"), as.Date("2021-12-01"), by = "month")
panel_grid <- expand_grid(
  buffer_id = unique(const_buffers$buffer_id),
  month = all_months
)

# Join buffer characteristics
panel_grid <- panel_grid %>%
  left_join(
    const_buffers %>%
      st_drop_geometry() %>%
      select(buffer_id, construction_id = ID, start_date, end_date, is_urban),
    by = "buffer_id"
  )

# Create treatment indicator
panel_grid <- panel_grid %>%
  mutate(
    treatment = month >= start_date & month <= end_date,
    post_treatment = month > end_date
  )

# Count accidents by buffer and month
monthly_accidents <- accidents_in_buffers %>%
  st_drop_geometry() %>%
  mutate(month = floor_date(accident_date, "month")) %>%
  count(buffer_id, month) %>%
  rename(accident_count = n)

# Join accident counts to panel
panel_data <- panel_grid %>%
  left_join(monthly_accidents, by = c("buffer_id", "month")) %>%
  replace_na(list(accident_count = 0))

# Join traffic volume data
panel_data <- panel_data %>%
  left_join(buffer_aadt, by = "buffer_id")

# Create normalized accident rate
panel_data <- panel_data %>%
  mutate(
    accident_rate = ifelse(traffic_volume > 0, 
                          accident_count / traffic_volume * 10000,  # Per 10,000 vehicles
                          NA)
  )

# Display the first few rows of the panel data
head(panel_data)
```

### Econometric Models

We implemented several model specifications to ensure robustness and to explore heterogeneity in the effects:

#### Base Difference-in-Differences Model

Our baseline model takes the form:

$$accident\_count_{it} = \alpha + \beta_1 treatment_{it} + \gamma_t + \delta_i + \varepsilon_{it}$$

Where:
- $accident\_count_{it}$ is the number of accidents in buffer $i$ at time $t$
- $treatment_{it}$ is a binary indicator for active construction
- $\gamma_t$ represents time fixed effects
- $\delta_i$ represents location fixed effects
- $\varepsilon_{it}$ is the error term

```{r base_model}
# Basic DiD model
did_model <- feols(
  accident_count ~ treatment | buffer_id + month,
  data = panel_data
)

# Display model summary
summary(did_model)
```

#### Heterogeneity Analysis

To examine differential effects between urban and rural areas, we extended the base model with an interaction term:

$$accident\_count_{it} = \alpha + \beta_1 treatment_{it} + \beta_2 (treatment_{it} \times urban_i) + \gamma_t + \delta_i + \varepsilon_{it}$$

```{r heterogeneity_model}
# DiD with urban/rural interaction
did_urban_model <- feols(
  accident_count ~ treatment * is_urban | buffer_id + month,
  data = panel_data %>% filter(!is.na(is_urban))
)

# Display model summary
summary(did_urban_model)
```

Contrary to our initial hypothesis, we found significant heterogeneous effects between urban and rural environments. In urban areas, construction is associated with a more pronounced increase in accident rates (interaction coefficient = 0.163, p < 0.01), while rural areas show a potentially different pattern, suggesting that traffic management challenges differ substantially by environment.

#### Normalized Accident Rate Model

To account for traffic volume differences, we also estimated models using normalized accident rates:

$$accident\_rate_{it} = \alpha + \beta_1 treatment_{it} + \gamma_t + \delta_i + \varepsilon_{it}$$

```{r rate_model}
# DiD with normalized accident rate
did_rate_model <- feols(
  accident_rate ~ treatment | buffer_id + month,
  data = panel_data %>% filter(!is.na(accident_rate))
)

# Display model summary
summary(did_rate_model)

# DiD with urban/rural heterogeneity and normalized rate
did_rate_urban_model <- feols(
  accident_rate ~ treatment * is_urban | buffer_id + month,
  data = panel_data %>% filter(!is.na(accident_rate), !is.na(is_urban))
)

# Display model summary
summary(did_rate_urban_model)
```

#### Spatial Econometric Models

To account for spatial dependence, we implemented spatial lag models:

```{r spatial_models}
# Create a sample data for spatial analysis (one month snapshot)
test_month <- as.Date("2021-06-01")
spdata <- panel_data %>%
  filter(month == test_month)

# Add spatial coordinates to the panel data
spdata <- spdata %>%
  left_join(
    const_buffers %>%
      st_centroid() %>%
      mutate(
        X = st_coordinates(.)[,1],
        Y = st_coordinates(.)[,2]
      ) %>%
      st_drop_geometry() %>%
      select(buffer_id, X, Y),
    by = "buffer_id"
  )

# Create spatial weights
spw_coords <- spdata %>%
  select(X, Y) %>%
  as.matrix()

# Create spatial weights - might need to adjust distance for your specific data
sp_nb <- spdep::dnearneigh(spw_coords, d1 = 0, d2 = 5000)
spw <- spdep::nb2listw(sp_nb, style = "W", zero.policy = TRUE)

# Spatial lag model
spatial_model <- spatialreg::lagsarlm(
  accident_count ~ treatment + is_urban + log(traffic_volume + 1),
  data = spdata,
  listw = spw,
  method = "Matrix",
  zero.policy = TRUE
)

# Display model summary
summary(spatial_model)
```

#### Spatial Spillover Analysis

To examine how construction impacts extend beyond the immediate construction zone, we implemented distance band models:

```{r spillover_model}
# Create distance bands from construction sites
panel_data <- panel_data %>%
  mutate(
    distance_band = cut(
      distance_to_nearest,
      breaks = c(0, 500, 1000, 2000, 5000, Inf),
      labels = c("0-500m", "500-1000m", "1-2km", "2-5km", ">5km")
    )
  )

# Test for spillover effects across distance bands
spillover_model <- feols(
  accident_count ~ i(distance_band, treatment, ref = ">5km") | buffer_id + month,
  data = panel_data %>% filter(!is.na(distance_band))
)

# Display model summary
summary(spillover_model)
```

Our spatial spillover analysis reveals nuanced effects extending beyond the immediate construction zone. Notably, areas 500m to 2km from construction sites demonstrate statistically significant increases in accident rates (coefficients ranging from 0.308 to 0.312, p < 0.001), suggesting that safety management strategies should extend well beyond the immediate construction boundaries.

#### Event Study Analysis

To test for parallel trends and examine dynamic effects, we implemented an event study approach:

```{r event_study}
# Create relative time variable
panel_data <- panel_data %>%
  mutate(
    rel_time = as.numeric(difftime(month, start_date, units = "days")) / 30,
    rel_time_cat = cut(rel_time, 
                       breaks = c(-Inf, -12, -6, -3, 0, 3, 6, 12, Inf),
                       labels = c("<-12m", "-12to-6m", "-6to-3m", "-3to0m", 
                                  "0to3m", "3to6m", "6to12m", ">12m"))
  )

# Event study model
event_study_model <- feols(
  accident_count ~ i(rel_time_cat, ref = "-3to0m") | buffer_id + month,
  data = panel_data %>% filter(!is.na(rel_time_cat))
)

# Display model summary
summary(event_study_model)
```

# Results

## Descriptive Statistics

Our analysis covered a large number of accidents and construction projects across California in 2021. The spatial distribution analysis included the urban-rural breakdown, which is critical for understanding the heterogeneous effects of construction on accident rates.

```{r summary_stats_table}
# Calculate actual summary statistics
summary_stats_df <- panel_data %>%
  filter(!is.na(is_urban)) %>%
  group_by(treatment, is_urban) %>%
  summarize(
    num_observations = n(),
    mean_accidents = mean(accident_count, na.rm = TRUE),
    median_accidents = median(accident_count, na.rm = TRUE),
    max_accidents = max(accident_count, na.rm = TRUE),
    mean_accident_rate = mean(accident_rate, na.rm = TRUE)
  ) %>%
  ungroup()

kableExtra::kable(summary_stats_df, 
                 col.names = c("Construction", "Urban", "Observations", "Mean Accidents", 
                               "Median Accidents", "Max Accidents", "Mean Accident Rate"),
                 caption = "Summary Statistics by Treatment Status and Urban/Rural Classification",
                 booktabs = TRUE,
                 digits = 3) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                           full_width = TRUE) %>%
  kableExtra::row_spec(0, bold = TRUE)
```

The summary statistics reveal important patterns in accident occurrence by construction status and urban/rural setting, providing context for our main regression analyses.

## Difference-in-Differences Results

Table 3 presents the results from our main difference-in-differences models.

```{r did_models_table}
# Compare DiD models
did_models_table <- etable(
  did_model, 
  did_urban_model, 
  did_rate_model, 
  did_rate_urban_model,
  headers = c("Base", "Urban Het.", "Rate", "Rate Urban"),
  digits = 3
)

# Print the comparison table
print(did_models_table)
```

The models show how active construction is associated with changes in accident counts, with notable differences between urban and rural areas. The urban heterogeneity model in particular reveals important differential impacts. When examining normalized accident rates (per 10,000 vehicles), the effects may change in magnitude or significance, suggesting traffic volume plays a complex role in the relationship.

## Spatial Lag Model Results

The spatial lag model confirms the presence of significant spatial autocorrelation, indicating that accident patterns exhibit strong spatial clustering. After accounting for this spatial dependence, we can examine construction effects:

```{r spatial_model_table}
# Extract actual spatial model results
spatial_sum <- summary(spatial_model)

# Create data frame with model results
coef_table <- as.data.frame(spatial_sum$Coef)
rho_row <- data.frame(
  Estimate = spatial_sum$rho,
  Std_Error = sqrt(spatial_sum$rho.se),
  z_value = spatial_sum$rho / sqrt(spatial_sum$rho.se),
  p_value = 2 * pnorm(abs(spatial_sum$rho / sqrt(spatial_sum$rho.se)), lower.tail = FALSE)
)

names(rho_row) <- names(coef_table)

# Combine into one table
spatial_model_df <- rbind(
  coef_table,
  rho_row
)

# Add variable names
spatial_model_df$Variable <- c(
  "Intercept", 
  "Treatment", 
  "Urban", 
  "log(Traffic Volume)",
  "Spatial Lag (ρ)"
)

# Format p-values
spatial_model_df$p_value_fmt <- ifelse(
  spatial_model_df$`Pr(>|z|)` < 0.001, 
  "<0.001", 
  sprintf("%.3f", spatial_model_df$`Pr(>|z|)`)
)

# Select and reorder columns
spatial_model_df <- spatial_model_df %>%
  select(Variable, Estimate, `Std. Error` = `Std. Error`, `z value`, `Pr(>|z|)` = p_value_fmt)

kableExtra::kable(spatial_model_df, 
                 col.names = c("Variable", "Estimate", "Std. Error", "z-value", "p-value"),
                 caption = "Spatial Lag Model Results",
                 booktabs = TRUE,
                 digits = 3) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),
                           full_width = FALSE) %>%
  kableExtra::row_spec(0, bold = TRUE) %>%
  kableExtra::add_footnote(sprintf("Model statistics: Log likelihood: %.1f, AIC: %.1f", 
                                  spatial_sum$LL, spatial_sum$AIC), 
                          notation = "none")
```

The spatial model confirms the effects of construction on accident counts even after controlling for spatial dependence, providing additional evidence for the relationship.

## Spatial Spillover Effects

Our spillover analysis examined how construction impacts extend to different distance bands from the construction site. Figure 1 presents these results.

```{r spillover_figure, fig.cap="Treatment Effects by Distance from Construction", fig.width=8, fig.height=5}
# Extract actual spillover model results
plot_spillover <- coefplot(spillover_model, 
          drop = "Intercept", 
          xlab = "Effect on Accident Count", 
          main = "Treatment Effect by Distance from Construction")

print(plot_spillover)
```

The spillover analysis reveals a pattern of effects at different distances from construction sites, with particularly notable impacts at specific distances. This suggests that construction effects aren't limited to the immediate construction zone.

## Event Study Results

The event study approach allows us to examine the dynamic effects of construction and test the parallel trends assumption.

```{r event_study_figure, fig.cap="Event Study: Effects Relative to Construction Start", fig.width=8, fig.height=5}
# Plot event study
plot_event_study <- coefplot(event_study_model,
                           xlab = "Time Relative to Construction Start",
                           ylab = "Effect on Accident Count",
                           main = "Event Study: Effect of Construction on Accidents")

print(plot_event_study)
```

The event study results show patterns of accident rates before and after construction begins, providing insights into temporal dynamics and supporting or challenging the parallel trends assumption required for causal interpretation of our DiD results.

## Accident Severity Analysis

We also examined how construction affects accident severity, classified by severity levels.

```{r severity_analysis}
# Analyze actual accident severity by construction period
severity_analysis <- accidents_in_buffers %>%
  st_drop_geometry() %>%
  filter(!is.na(time_period)) %>%
  count(time_period, Severity) %>%
  group_by(time_period) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup()

# Format severity for better display
severity_analysis <- severity_analysis %>%
  mutate(
    Severity_factor = factor(
      Severity, 
      levels = sort(unique(Severity)),
      labels = c("Slight", "Moderate", "Serious", "Fatal")[1:length(unique(Severity))]
    )
  )

# Display severity distribution
kable(
  severity_analysis %>%
    select(time_period, Severity, n, proportion) %>%
    mutate(proportion = proportion * 100) %>%
    arrange(time_period, Severity),
  col.names = c("Construction Period", "Severity Level", "Count", "Percentage (%)"),
  caption = "Distribution of Accident Severity by Construction Period",
  digits = 1
)
```

```{r severity_figure, fig.cap="Accident Severity by Construction Period", fig.width=8, fig.height=5}
# Visualize accident severity distribution
ggplot(severity_analysis, aes(x = time_period, y = proportion, fill = Severity_factor)) +
  geom_col() +
  scale_fill_brewer(palette = "Blues") +
  labs(
    title = "Accident Severity Distribution by Construction Period",
    x = "Relation to Construction",
    y = "Proportion of Accidents",
    fill = "Severity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(size = 12),
    panel.grid.minor = element_blank()
  )
```

The severity analysis reveals a consistent distribution of accident severity across construction periods. Approximately 87% of accidents are classified as moderate severity, with about 12% categorized as serious. Interestingly, the proportional stability suggests that while construction impacts accident frequency, it does not substantially alter the underlying severity distribution.

# Discussion

## Interpretation of Main Findings

Our analysis provides robust evidence that road construction causally impacts traffic accident rates, with important heterogeneity across urban and rural environments. The main findings can be summarized as follows:

1. **Overall Construction Effect**: Our base difference-in-differences model reveals a statistically significant increase in accident counts during construction periods (coefficient = 0.101, p < 0.001), indicating that road construction activities are associated with approximately 10.1% more accidents compared to non-construction periods..

2. **Urban-Rural Heterogeneity**: The impact of construction on accidents differs substantially between urban and rural environments. In urban areas, construction is generally associated with increased accident counts, while in rural areas, there is evidence of potentially reduced accidents during construction. This heterogeneity may be explained by several factors:
   - Higher traffic density in urban areas creates more complex interactions around construction zones
   - Rural construction may induce more cautious driving or effective traffic diversion
   - Urban construction often involves more complex traffic management challenges

3. **Spatial Spillover Effects**: The spatial spillover analysis shows that construction impacts extend beyond the immediate construction zone, with effects varying by distance. This finding has important implications for traffic management strategies, suggesting that safety measures should extend beyond the construction zone itself to areas where merging and speed adjustment typically occur.

4. **Temporal Dynamics**: The event study results provide insights into accident patterns before, during, and after construction. We observe changes in accident rates during different phases of construction, which may reflect adaptation effects as drivers adjust to changed road conditions.

5. **Severity Patterns**: Our severity analysis shows the distribution of accident severity levels (Slight, Moderate, Serious, Fatal) across different construction periods. This helps us understand whether construction not only affects accident frequency but also severity levels.

## Policy Implications

These findings have several important policy implications for transportation safety planning:

1. **Targeted Resource Allocation**: Traffic management resources should be allocated differentially between urban and rural construction zones, with greater emphasis on urban areas where construction effects on accidents are typically larger.

2. **Extended Safety Zones**: Traffic safety measures should not be limited to the immediate construction zone but should extend to areas surrounding construction sites where our spillover analysis indicates significant effects.

3. **Temporal Planning**: The temporal patterns revealed in our event study suggest that enhanced safety measures may be particularly important during specific phases of construction projects.

4. **Severity-Focused Strategies**: Our severity analysis provides insights for emergency response planning and the design of safety measures aimed at mitigating more severe accidents.

## Limitations and Future Research

While our study provides evidence on the causal impact of construction on accidents, several limitations should be noted:

1. **Time Period**: Our analysis is limited to data from 2021, which may have been affected by pandemic-related traffic patterns. Future research should examine longer time periods and potential yearly variations.

2. **Granularity**: While we distinguish between urban and rural areas, more detailed classifications (e.g., suburban, exurban) could provide additional insights for targeted interventions.

3. **Construction Types**: Our data does not allow us to consistently differentiate between different types of construction activities (e.g., resurfacing, bridge repair, lane expansion), which may have different safety implications.

4. **Weather and Seasonal Effects**: While our models include time fixed effects that partly control for seasonal patterns, more explicit modeling of weather interactions with construction effects could yield valuable insights.

5. **Endogeneity**: Despite our use of a DiD framework, unobserved confounders may still bias our estimates. Future research could explore instrumental variable approaches or quasi-experimental designs to address endogeneity concerns.

6. **AADT data**: The AADT data used for traffic volume normalization may include estimates or projections that do not precisely match the 2021 study period, potentially introducing minor measurement uncertainties."

Future research should address these limitations and further explore the mechanisms underlying the observed heterogeneity in construction effects.

# Conclusion

This study provides evidence on the relationship between road construction and traffic accidents in California. Using high-resolution spatial data and advanced econometric methods, we demonstrate that construction activities significantly impact accident rates, with effects that vary substantially between urban and rural environments and extend beyond the immediate construction zone.

The findings have important implications for transportation safety planning, suggesting the need for targeted resource allocation, extended safety zones, and temporally-focused safety measures. By implementing these recommendations, transportation authorities can better mitigate the safety risks associated with necessary infrastructure maintenance and development.

Our methodological approach combines spatial analysis with causal inference techniques, providing a framework that can be extended to other regions and contexts. By continuing to refine these methods and addressing the limitations noted, future research can further enhance our understanding of the complex relationship between construction activities and traffic safety.

# References

Ahmed, M. M., Franke, R., Ksaibati, K., & Shinstine, D. S. (2018). Effects of truck traffic on crash injury severity on rural highways in Wyoming using Bayesian binary logit models. *Accident Analysis & Prevention*, 117, 106-113.

Li, Y., & Bai, Y. (2008). Comparison of characteristics between fatal and injury accidents in the highway construction zones. *Safety Science*, 46(4), 646-660.